### 2017-02-07 doc/Paper.tex
Create Paper.tex

\documentclass[a4paper]{article}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{graphicx}  % graphics with figure-environment
\usepackage{hyperref}
% colors for hyperlinks
% colored borders (false) colored text (true)
\hypersetup{colorlinks=true,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black}
\usepackage[authoryear,round]{natbib} % package for bibliography
% package for header
\usepackage[automark]{scrpage2}
\pagestyle{scrheadings}
\ihead[]{Handwritten character recognition in MPI}
\ohead[]{\today}
\cfoot[]{\pagemark} 
\setheadsepline[122mm]{0.3mm}
\begin{document}
	\title{
	\Huge Handwritten digit recognition
	}

	\vspace{2cm}

	\author{\Large \href{mailto:stefan.niculae@my.fmi.unibuc.ro}{Stefan Niculae} \and \Large \href{mailto:ionut.ciocoiu@my.fmi.unibuc.ro}{Ionut Ciocoiu}
	\vspace{3cm}}

	\date{
	\large Parallel \& Concurrent Programming Lab Report \\
    \vspace{0.2cm}
	\today
	}

	\maketitle
	\setlength{\parindent}{0pt}

\vspace{5cm}
\begin{abstract}
We implement a neural network to classify handwritten digits. The training is done in parallel, using MPI. We provide a comparison for model hyper-parameters and for training time vs number of processes. It achieves great accuracy and fast training time. An interactive demo is also presented

\end{abstract}
	\newpage
	\tableofcontents
	\newpage

\section{Problem statement} % (fold)
\label{sec:problem}
Given a picture of a handwritten digit, label it accordingly 0 - 9.

\section{Section about quotations} % (fold)
\label{sec:section_about_quotations}

In this section, an example for a literal quotation is given.

\begin{quotation}
	\emph{``A persona is a rich picture of an imaginary person who represents your core user group.''}
	\citep{Dix04}
\end{quotation}

Sometimes you might want to make use of the authors name within the text. Before, we used the command \texttt{citep\{\}}, which creates the brackets around author name and year. You can also use the \texttt{cite} command like this: \\

\cite{Dix04} defined the concept of persona as follows: 
\begin{quotation}
	\emph{``A persona is a rich picture of an imaginary person who represents your core user group.''}
	\citep{Dix04}
\end{quotation}

You may notice, that this increases the readability of the text. \\

According to APA format\footnote{ American Psychological Association (APA)} there are some rules, when and how to include page numbers, when referring to literature.

\begin{quotation}
	\emph{``Include page numbers for any citations in the text of your paper that include direct quotations or refer to a specific part of the work you are referencing. Direct quotations must include a page number as part of the citation. The quoted material should be followed by a citation in parentheses that gives the author's name, the year in which the work was published, and the page number from which the quoted material appears.''}
	\citep{Hall}
\end{quotation}

Check out the example and recommendations of \cite{Hall} on \url{http://www.ehow.com/how_5689799_cite-numbers-apa-format.html}. In \LaTeX you can include the pages very easy. For example: \\

\citet[p. 86]{Baddeley:1974ts} stated:

\begin{quotation}
	\emph{``We hope that our preliminary attempts to begin answering the question will convince the reader, not necessarily that our views are correct, but that the question was and is well worth asking''}
	\citep[p. 86]{Baddeley:1974ts}
\end{quotation}

Note that in the first reference, we used \texttt{citet[]\{\}} in order to have brackets just around year and page number; later we used \texttt{citep[]\{\}}.

% section section_about_quotations (end)

\section{Section about references within the document} % (fold)
\label{sec:section_about_references_within_the_document}

If you want to refer to you own chapters, figures, tables or the like, you can make use of the \texttt{ref\{\}} command, for example:
\begin{itemize}
	\item section~\ref{sec:section_about_quotations} on page \pageref{sec:section_about_quotations}
\end{itemize}

% section section_about_references_within_the_document (end)

\subsection{Subsection within Foundations} % (fold)
\label{sub:subsection_within_foundations}
Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
% subsection subsection_within_foundations (end)

\subsection{Another subsection within Foundations} % (fold)
\label{sub:another_subsection_within_foundations}
Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
% subsection another_subsection_within_foundations (end)

% section foundations (end)

\section{Methodology} % (fold)
\label{sec:methodology}
Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua \citep{Dix04}. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
% section methodology (end)

\section{Results} % (fold)
\label{sec:results}
Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua \citep[p. 48]{Baddeley:1974ts}. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
% section results (end)

\section{Conclusion} % (fold)
\label{sec:conclusion}
Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
% section conclusion (end)

\newpage

\bibliographystyle{natdin}
	\bibliography{references} % expects file "references.bib"
	\addcontentsline{toc}{section}{References}
\end{document}

### 2017-02-07 doc/Paper.tex
Add bibliography & add dataset section

@@
 \hypersetup{colorlinks=true,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black}
-\usepackage[authoryear,round]{natbib} % package for bibliography
+\usepackage{cite}  % bibliograby
 % package for header
@@
 
-\section{Problem statement} % (fold)
+\section{Problem statement}
 \label{sec:problem}
@@
 
-\section{Section about quotations} % (fold)
-\label{sec:section_about_quotations}
+\section{Dataset}
+\label{sec:dataset}
+We demonstrate our model performance on the MNIST database, widely used dataset for training and testing in the field of machine learning.
+It contains 70 000 examples of labeled handwritten digits.
+The digits have been size-normalized and centered in a fixed-size image \cite{mnist}.
 
-In this section, an example for a literal quotation is given.
+\begin{figure}[htb]
+\centering
+\includegraphics[width=8cm]{images/mnist-images.png}
+\caption{Sample MNIST images}
+\end{figure}
 
-\begin{quotation}
-	\emph{``A persona is a rich picture of an imaginary person who represents your core user group.''}
-	\citep{Dix04}
-\end{quotation}
+The images are black-and-white, with a resolution of $28 \times 28$ pixels, represented as an array of $784$ values. Each value ranges from $0$ to $1$ indicating the amount of blackness in the pixel.
 
-Sometimes you might want to make use of the authors name within the text.
-Before, we used the command \texttt{citep\{\}}, which creates the brackets around author name and year.
-You can also use the \texttt{cite} command like this: \\
-
-\cite{Dix04} defined the concept of persona as follows: 
-\begin{quotation}
-	\emph{``A persona is a rich picture of an imaginary person who represents your core user group.''}
-	\citep{Dix04}
-\end{quotation}
-
-You may notice, that this increases the readability of the text. \\
-
-According to APA format\footnote{ American Psychological Association (APA)} there are some rules, when and how to include page numbers, when referring to literature.
-
-\begin{quotation}
-	\emph{``Include page numbers for any citations in the text of your paper that include direct quotations or refer to a specific part of the work you are referencing.
-Direct quotations must include a page number as part of the citation.
-The quoted material should be followed by a citation in parentheses that gives the author's name, the year in which the work was published, and the page number from which the quoted material appears.''}
-	\citep{Hall}
-\end{quotation}
-
-Check out the example and recommendations of \cite{Hall} on \url{http://www.ehow.com/how_5689799_cite-numbers-apa-format.html}. In \LaTeX you can include the pages very easy. For example: \\
-
-\citet[p. 86]{Baddeley:1974ts} stated:
-
-\begin{quotation}
-	\emph{``We hope that our preliminary attempts to begin answering the question will convince the reader, not necessarily that our views are correct, but that the question was and is well worth asking''}
-	\citep[p. 86]{Baddeley:1974ts}
-\end{quotation}
-
-Note that in the first reference, we used \texttt{citet[]\{\}} in order to have brackets just around year and page number; later we used \texttt{citep[]\{\}}.
-
-% section section_about_quotations (end)
-
-\section{Section about references within the document} % (fold)
-\label{sec:section_about_references_within_the_document}
-
-If you want to refer to you own chapters, figures, tables or the like, you can make use of the \texttt{ref\{\}} command, for example:
-\begin{itemize}
-	\item section~\ref{sec:section_about_quotations} on page \pageref{sec:section_about_quotations}
-\end{itemize}
-
-% section section_about_references_within_the_document (end)
-
-\subsection{Subsection within Foundations} % (fold)
-\label{sub:subsection_within_foundations}
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-% subsection subsection_within_foundations (end)
-
-\subsection{Another subsection within Foundations} % (fold)
-\label{sub:another_subsection_within_foundations}
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-% subsection another_subsection_within_foundations (end)
-
-% section foundations (end)
-
-\section{Methodology} % (fold)
-\label{sec:methodology}
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua \citep{Dix04}.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-% section methodology (end)
-
-\section{Results} % (fold)
-\label{sec:results}
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua \citep[p.
-48]{Baddeley:1974ts}.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-% section results (end)
-
-\section{Conclusion} % (fold)
-\label{sec:conclusion}
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-% section conclusion (end)
+\begin{figure}[htb]
+\centering
+\includegraphics[width=8cm]{images/mnist-matrix.png}
+\caption{Matrix representation of a digit}
+\end{figure}
 
@@
 
-\bibliographystyle{natdin}
+\bibliographystyle{ieeetr}
 	\bibliography{references} % expects file "references.bib"

### 2017-02-07 doc/Paper.tex
Add model description, plot prototype

@@
 \documentclass[a4paper]{article}
+
 \usepackage[english]{babel}
 \usepackage[utf8x]{inputenc}
-\usepackage{graphicx}  % graphics with figure-environment
-\usepackage{hyperref}
-% colors for hyperlinks
-% colored borders (false) colored text (true)
-\hypersetup{colorlinks=true,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black}
+
 \usepackage{cite}  % bibliograby
-% package for header
-\usepackage[automark]{scrpage2}
+\usepackage{graphicx}  % images
+\usepackage{pgfplots}  % plots
+  \pgfplotsset{width=10cm,compat=1.9} % plots settings
+  \usepgfplotslibrary{external}
+  \tikzexternalize
+
+\usepackage{hyperref}  % references
+  \hypersetup{colorlinks=true,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black} % references settings
+
+\usepackage[automark]{scrpage2} % heading
   \pagestyle{scrheadings}
-\ihead[]{Handwritten character recognition in MPI}
-\ohead[]{\today}
-\cfoot[]{\pagemark} 
+  \ihead[]{Handwritten character recognition in MPI} % left header
+  \ohead[]{\today} % right header
+  \cfoot[]{\pagemark} % footer
   \setheadsepline[122mm]{0.3mm}
+
 \begin{document}
@@
 An interactive demo is also presented
+\end{abstract}
 
-\end{abstract}
 \newpage
 \tableofcontents
+
 \newpage
-
 \section{Problem statement}
@@
 \includegraphics[width=8cm]{images/mnist-images.png}
-\caption{Sample MNIST images}
+\caption{Sample MNIST images \cite{tf}}
 \end{figure}
@@
 \includegraphics[width=8cm]{images/mnist-matrix.png}
-\caption{Matrix representation of a digit}
+\caption{Matrix representation of a digit \cite{tf}}
 \end{figure}
@@
 \newpage
+\section{Model}
+\label{sec:model}
+We use a back-propagation neural network with cross-entropy as cost function and soft-max activation.
+L2 regularization is used to limit overfitting and adaptive learning rate helps soften instability in later epochs.
 
+\begin{figure}[htb]
+\centering
+\includegraphics[width=8cm]{images/model-graph.png}
+\caption{Simplified representation of model graph \cite{tf}}
+\end{figure}
+
+\paragraph{Training} We employ gradient-descent on mini-batches. Early-stopping prevents the model from further learning when performance on the validation split starts decreasing.
+
+\paragraph{Evaluation}
+TODO
+
+\begin{tikzpicture}
+\centering
+\begin{axis} [
+  title={Training performance},
+  xlabel={Epoch},
+  ylabel={Accuracy (\%)},
+  xmin=0, xmax=100,
+  ymin=0, ymax=100,
+  xtick={0,25,50,75,100},
+  ytick={0,25,50,75,100},
+  legend pos=south east,
+]
+
+\addplot[]
+  coordinates {
+  (0,25)(25,65)(50,90)(100,95)
+  };
+\addplot[dashed]
+  coordinates {
+  (0,25)(25,50)(50,80)(100,90)
+  };
+
+\legend{train set, test set}
+\end{axis}
+\end{tikzpicture}
+
+\newpage
 \bibliographystyle{ieeetr}
@@
 	\addcontentsline{toc}{section}{References}
+
 \end{document}

### 2017-02-07 doc/Paper.tex
Add model description

@@
 
+\usepackage{amsmath}  % equations
 \usepackage{cite}  % bibliograby
@@
 \label{sec:model}
-We use a back-propagation neural network with cross-entropy as cost function and soft-max activation.
-L2 regularization is used to limit overfitting and adaptive learning rate helps soften instability in later epochs.
+We use a back-propagation neural network:
+\begin{align*}
+W^{(k)} &= W^{(k-1)} + \alpha \nabla_W^{(k-1)} \\ 
+b^{(k)} &= b^{(k-1)} + \alpha \nabla_b^{(k-1)}
+\end{align*}
+Where $W$ is the weights matrix, $10 \times 784$ -- the number of classes( one for each digit) and the dimension of an image ($28 \times 28$).
+$b$ is the bias vector, $1 \times 10$, corresponding to the free element.
+$W^{(k)}$ denotes the weights at step $k$.
+$\nabla$ represents the update which must produce to improve prediction.
+\\
+
+The activation function is softmax:
+$$f(x_i)=\frac{exp(x_i)}{\sum_j exp(x_j)}$$
+Where $exp(x)$ represents $e^x$, Euler's constant. It \textit{normalizes} the vector, causing the greatest elements to have logarithmic relative value.
+\\
+
+The cost function is cross-entropy:
+$$H(y_t, y_p)=-\sum_x y_t log(y_p)$$
+Where $y_t$ is the true, correct label of an image and $y_p$ is the label predicted by the model.
+\\
+
+$L_2$ regularization is used:
+$$\frac{1}{\lambda} \sum_i \sum_j W_{i,j}^2$$
+Where $\lambda$ is the regularization parameter. This limits overfitting by imposing a penalty on big weights.
+\\
+
+The model uses an adaptive learning rate:
+$$\eta \frac{ W^{(k)} - W^{(k-1)} } { \nabla^{(k)} - \nabla^{(k-1)} }$$
+Where $\eta$ is the momentum parameter. This helps with stability in late epochs.
 

### 2017-02-07 doc/Paper.tex
Add Implementation section & TODOs

@@
 \begin{abstract}
+TODO: rewrite when paper is done, talk about each section
+
 We implement a neural network to classify handwritten digits.
@@
 
-\newpage
-\tableofcontents
+% \newpage
+% \pagestyle{plain}
+% \tableofcontents
 
 \newpage
+TODO: introduction?
+
 \section{Problem statement}
@@
 \label{sec:model}
+In order to recognize a given image, we need to model to understand the data. It needs to be able to be queried for a new instance in order to provide an interactive demo.
+\\
+
 We use a back-propagation neural network:
@@
 
+\paragraph{Prediction}
+TODO
+
 \paragraph{Evaluation}
@@
 \newpage
+\section{Parallelization}
+Parallelism is needed in the training phase of the model, as prediction happens instantaneously
+
+TODO: talk about matrix multiplication overhead
+
+TODO: graph of train time vs nr of processors
+
+\newpage
+\section{Graphical Interface}
+TODO: picture of resizing
+
+TODO: picture of GUI
+
+\newpage
+\section{Implementation}
+After going through the high-level overview of the project, we now delve into implementation details.
+
+\paragraph{Predictive model} The core of the project is implemented in C++.
+This includes the neural network training and prediction.
+Data is maintained in memory using vectors from the standard library.
+
+\paragraph{Graphical Interface} The GUI is implemented as a web page in HTML, with a canvas for the input box.
+The drawing is facilitated by the Sketch library \cite{sketch}. Interactivity and picture resizing is done through JavaScript.
+
+\paragraph{Communication Server} The glue between the front-end and the model is a web-server done in Python, using the Flask library \cite{flask}.
+An array of $784$ pixels is received through from the GUI and is passed to the compiled executable.
+The output (predicted digit and prediction confidence) is then passed back through AJAX.
+\\
+
+The entire source code is available at \url{https://github.com/stefan1niculae/mnist-in-mpi}.
+
+\vspace{\fill} % place references at the bottom
 \bibliographystyle{ieeetr}

### 2017-02-07 main.cpp
mpi with row-by-row communication

@@ -72,4 +72,3 @@ int main(int argc, const char* argv[]) {
         return 0;
     }
 }
-

### 2017-02-07 matrix.cpp

@@ -125,7 +125,7 @@ void add(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
     if ((lhs.n_rows != rhs.n_rows) || (lhs.n_cols != rhs.n_cols))
         throw runtime_error(string_format("Matrix addition: number of rows/cols is different: "
                                                   "lhs = (%d, %d), rhs = (%d, %d)",
-        lhs.n_rows, lhs.n_cols, rhs.n_rows, rhs.n_cols));
+                                          lhs.n_rows, lhs.n_cols, rhs.n_rows, rhs.n_cols));
 
     for (int i = 0; i < rhs.n_rows; ++i)
         for (int j = 0; j < rhs.n_cols; ++j)
@@ -193,7 +193,7 @@ void scalar_mult(double scalar, const Matrix& matrix, Matrix& result) {
 void dot(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
     if (lhs.n_cols != rhs.n_rows)
         throw runtime_error(string_format("Dimensions do not agree for matrix multiplication: "
-                      "lhs cols = %d, rhs rows = %d", lhs.n_cols, rhs.n_rows));
+                                                  "lhs cols = %d, rhs rows = %d", lhs.n_cols, rhs.n_rows));
 
     int n = lhs.n_rows;
     int p = lhs.n_cols;

### 2017-02-07 main.cpp
mpi with row-by-row communication

@@ -1,8 +1,10 @@
 #include <iostream>
 #include "reading.cpp"
 #include "nn.cpp"
+#include <iomanip>
 
 using namespace std;
+
 
 void train_and_save() {
     // TODO: names as additional parameters
@@ -31,7 +33,15 @@ int predict(string pixels_string) {
     return digit_predicted;
 }
 
-int main(int argc, const char* argv[]) {
+int main(int argc, char* argv[]) {
+    Matrix images, labels;
+    read_data("/Users/Stefan/Projects/mnist-in-mpi/data/train.csv", images, labels, true);
+    NeuralNetwork net;
+    vector<double> costs, accs;
+    cout << setprecision(3);
+    net.train(images, labels, costs, accs, 35, 400, .1);
+    return 0;
+
     if (argc < 2) {
         cout << "usage: mnist_in_mpi.out (train|predict)" << endl;
         return 1;

### 2017-02-07 matrix.cpp

@@ -42,6 +42,11 @@ inline int n_cols(const vector<vector<T>>& m) {
 }
 
 template <class T>
+inline int n_elements(const vector<vector<T>>& m) {
+    return n_rows(m) * n_cols(m);
+}
+
+template <class T>
 vector<T> chunk(const vector<T>& m, int from, int to) {
     if (from < 0)
         throw "From < 0";
@@ -73,6 +78,12 @@ vector<vector<T>> col_wise_sums(const vector<vector<T>>& matrix) {
 
 
 /*** Operations ***/
+void clear(Matrix& m) {
+    for (auto& row : m)
+        for (auto& elem : row)
+            elem = 0.;
+}
+
 
 // matrix transposition
 template <class T>
@@ -149,6 +160,12 @@ vector<vector<T>> operator+ (const vector<vector<T>>& lhs, const vector<vector<T
         for (int j = 0; j < n_cols(rhs); ++j)
             result[i][j] += rhs[i][j];
     return result;
+}
+
+template <class T>
+vector<vector<T>> operator+= (vector<vector<T>>& lhs, const vector<vector<T>>& rhs) {
+    lhs = lhs + rhs;
+    return lhs;
 }
 
 // matrix - vector addition

### 2017-02-07 nn.cpp

@@ -5,9 +5,18 @@
 #include <fstream>
 #include "matrix.cpp"
 #include "evaluate.cpp"
+#include "mpi.h"
 
 
 using namespace std;
+
+
+/*** MPI constants ***/
+const int MASTER = 0;
+const int GRAD_W_TAG    = 0;
+const int GRAD_B_TAG    = 1;
+//const int UPDATED_W_TAG = 2;
+
 
 
 /*** Requisites ***/
@@ -96,26 +105,104 @@ public:
                vector<double>& cost_history, vector<double>& accuracy_history,
                int n_epochs=10, int batch_size=100, double lr=0.1,
                bool verbose=true) {
+        // MPI initialization
+        MPI_Init(NULL, NULL);
+        int rank, n_processes;
+        MPI_Comm_rank(MPI_COMM_WORLD, &rank);
+        MPI_Comm_size(MPI_COMM_WORLD, &n_processes);
+        int n_workers = n_processes - 1; // first process is the master
+        int chunk_size = batch_size / n_workers;
+
+
         double cost, acc;
         auto Y_labels = from_one_hot_matrix(Y); // {5, 2, 9, ... }
 
-        Matrix grad_W = blank_matrix(data_dim, n_classes, 0.);
-        Matrix grad_b = blank_matrix(1, n_classes, 0.);
+        Matrix total_grad_W = blank_matrix(data_dim, n_classes, 0.);
+        Matrix total_grad_b = blank_matrix(1, n_classes, 0.);
+        Matrix partial_grad_W = blank_matrix(data_dim, n_classes, 0.);
+        Matrix partial_grad_b = blank_matrix(1, n_classes, 0.);
 
         for (int epoch = 1; epoch <= n_epochs; ++epoch) {
-            if (verbose)
+            if (verbose && rank == 0)
                 cout << string_format("Epoch %d / %d", epoch, n_epochs) << endl;
 
             for (int i = 0; i < n_rows(X); i+=batch_size) {
                 // TODO: make random batch generator
-                cost = grad(chunk(X, i, i+batch_size), chunk(Y, i, i+batch_size),
-                            grad_W,  grad_b);
+                if (rank == 0) { // master
+//                    cout << "master: will wait for workers for batch " << i << " to " << i+batch_size << endl;
+                    // Clear the total gradient from previous communications
+                    clear(total_grad_W);
+                    clear(total_grad_b);
 
-                // TODO regularization
-                W = W - lr * grad_W;
[…]

### 2017-02-07 reading.cpp

@@ -11,7 +11,7 @@ using namespace std;
 
 const int PIXEL_DIM = 255;
 
-void read_data(string filename, Matrix& images, Matrix& labels) {
+void read_data(string filename, Matrix& images, Matrix& labels, bool verbose= true) {
     ifstream file(filename);
 
     if (!file.is_open())
@@ -41,7 +41,8 @@ void read_data(string filename, Matrix& images, Matrix& labels) {
     for (auto val : label_values)
         labels.push_back(make_one_hot(val));
 
-    cout << "Done reading from " + filename << endl;
+    if (verbose)
+        cout << "Done reading from " + filename << endl;
 }
 
 void print_image(const vector<double>& pixels) {

### 2017-02-07 matrix.cpp
Add Matrix class

@@ -1,228 +1,261 @@
 #pragma once
 
-#include <vector>
-#include <algorithm>  // max_element
 #include <math.h>  // exp, math
 #include "utils.cpp"
 
-using namespace std;
 
-typedef vector<vector<double>> Matrix;
+class Matrix {
 
+public:
+    double** data;
+    int n_rows;
+    int n_cols;
 
-template <class T>
-ostream& operator<<(ostream& os, const vector<vector<T>>& matrix) {
-    // Even though the << operator can apply twice for <vector<vector>>, this way we can add endlines
-    for (const auto& row : matrix)
-        os << row << endl;
-    return os;
-}
-
-template <class T>
-istream& operator>>(istream& is, vector<vector<T>>& matrix) {
-    // Read until end of stream
-    vector<T> v;
-    while (is >> v) {
-        matrix.push_back(v);
-        v.clear();
+    Matrix(int n_rows,int n_cols): n_rows(n_rows), n_cols(n_cols) {
+        data = new double*[n_rows];
+        for (int i = 0; i < n_rows; ++i)
+            data[i] = new double[n_cols];
     }
-    return is;
-}
 
-template <class T>
-inline int n_rows(const vector<vector<T>>& m) {
-    return m.size();
-}
-
-template <class T>
-inline int n_cols(const vector<vector<T>>& m) {
-    return n_rows(m) == 0 ?
-           0 :
-           m[0].size();
-}
-
-template <class T>
-vector<T> chunk(const vector<T>& m, int from, int to) {
-    if (from < 0)
-        throw "From < 0";
-    if (to > m.size()) to = m.size();
-    return vector<T>(&m[from], &m[to]);
-}
-
-template <class T>
-inline vector<vector<T>> blank_matrix(int rows, int cols) {
-    // Default value
-    return vector<vector<T>>(rows, vector<T>(cols));
-}
-template <class T>
-inline vector<vector<T>> blank_matrix(int rows, int cols, T value) {
-    return vector<vector<T>>(rows, vector<T>(cols, value));
-}
-
-template <class T>
-vector<vector<T>> col_wise_sums(const vector<vector<T>>& matrix) {
-    vector<T> sums(n_cols(matrix), 0.);
-    for (int r = 0; r < n_rows(matrix); ++r)
-        for (int c = 0; c < n_cols(matrix); ++c)
-            sums[c] += matrix[r][c];
-
-    Matrix result = {sums}; // make it a 1xn matrix
-    return result;
-}
-
-
-
-/*** Operations ***/
-
[…]

### 2017-02-07 matrix.cpp
Add IO

@@ -22,11 +22,31 @@ public:
             delete[] data[i];
         delete[] data;
     }
-
 };
 
-/*** Operators ***/
 
+
+/*** IO ***/
+ostream& operator<<(ostream& os, const Matrix& matrix) {
+    for (int i = 0; i < matrix.n_rows; ++i) {
+        for (int j = 0; j < matrix.n_cols; ++j)
+            os << matrix[i][j] << ' ';
+        os << endl;
+    }
+    return os;
+}
+istream& operator>>(istream& is, Matrix& matrix) {
+    for (int i = 0; i < matrix.n_rows; ++i)
+        for (int j = 0; j < matrix.n_cols; ++j)
+            is >> matrix[i][j];
+    return is;
+}
+
+
+
+
+
+/*** Operators ***/
 void add_to_each(Matrix& matrix, const Matrix& to_add) {
     if (matrix.n_cols != to_add.n_cols)
         throw runtime_error(string_format("Add to each: matrix width and vector size are different: "

### 2017-02-07 matrix.cpp
Add clear to Matrix

@@ -22,6 +22,12 @@ public:
             delete[] data[i];
         delete[] data;
     }
+
+    clear() {
+        for (int i = 0; i < n_rows; ++i)
+            for (int j = 0; j < n_cols; ++j)
+                data[i][j] = 0;
+    }
 };
 
 
@@ -62,54 +68,12 @@ void add_to_each(Matrix& matrix, const Matrix& to_add) {
 
 
 
-//
-//template <class T>
-//ostream& operator<<(ostream& os, const vector<vector<T>>& matrix) {
-//    // Even though the << operator can apply twice for <vector<vector>>, this way we can add endlines
-//    for (const auto& row : matrix)
-//        os << row << endl;
-//    return os;
-//}
-//
-//template <class T>
-//istream& operator>>(istream& is, vector<vector<T>>& matrix) {
-//    // Read until end of stream
-//    vector<T> v;
-//    while (is >> v) {
-//        matrix.push_back(v);
-//        v.clear();
-//    }
-//    return is;
-//}
-//
-//template <class T>
-//inline int n_rows(const vector<vector<T>>& m) {
-//    return m.size();
-//}
-//
-//template <class T>
-//inline int n_cols(const vector<vector<T>>& m) {
-//    return n_rows(m) == 0 ?
-//           0 :
-//           m[0].size();
-//}
-//
 //template <class T>
 //vector<T> chunk(const vector<T>& m, int from, int to) {
 //    if (from < 0)
 //        throw "From < 0";
 //    if (to > m.size()) to = m.size();
 //    return vector<T>(&m[from], &m[to]);
-//}
-//
-//template <class T>
-//inline vector<vector<T>> blank_matrix(int rows, int cols) {
-//    // Default value
-//    return vector<vector<T>>(rows, vector<T>(cols));
-//}
-//template <class T>
-//inline vector<vector<T>> blank_matrix(int rows, int cols, T value) {
-//    return vector<vector<T>>(rows, vector<T>(cols, value));
 //}
 //
 //template <class T>

### 2017-02-07 matrix.cpp
Add Col-wise-sum

@@ -48,10 +48,6 @@ istream& operator>>(istream& is, Matrix& matrix) {
     return is;
 }
 
-
-
-
-
 /*** Operators ***/
 void add_to_each(Matrix& matrix, const Matrix& to_add) {
     if (matrix.n_cols != to_add.n_cols)
@@ -66,26 +62,24 @@ void add_to_each(Matrix& matrix, const Matrix& to_add) {
             matrix.data[r][c] += to_add.data[0][c];
 }
 
+void col_wise_sums(const Matrix& matrix, Matrix& result) {
+    result.clear();
+    for (int r = 0; r < matrix.n_rows; ++r) {
+        for (int c = 0; c < matrix.n_cols; ++c)
+            result[c] += matrix[r][c];
+    }
+}
 
 
-//template <class T>
-//vector<T> chunk(const vector<T>& m, int from, int to) {
+//Matrix chunk(const vector<T>& m, int from, int to) {
 //    if (from < 0)
 //        throw "From < 0";
 //    if (to > m.size()) to = m.size();
 //    return vector<T>(&m[from], &m[to]);
 //}
-//
-//template <class T>
-//vector<vector<T>> col_wise_sums(const vector<vector<T>>& matrix) {
-//    vector<T> sums(n_cols(matrix), 0.);
-//    for (int r = 0; r < n_rows(matrix); ++r)
-//        for (int c = 0; c < n_cols(matrix); ++c)
-//            sums[c] += matrix[r][c];
-//
-//    Matrix result = {sums}; // make it a 1xn matrix
-//    return result;
-//}
+
+
+
 //
 //
 //

### 2017-02-07 matrix.cpp
Add transpose

@@ -86,17 +86,14 @@ void col_wise_sums(const Matrix& matrix, Matrix& result) {
 ///*** Operations ***/
 //
 //// matrix transposition
-//template <class T>
-//vector<vector<T>> transpose(const vector<vector<T>>& matrix) {
-//    vector<vector<T>> transposed = vector<vector<T>> (n_cols(matrix), vector<T>(n_rows(matrix)));
-//    for (int j=0; j<n_cols(matrix); j++) {
-//        for (int i=0; i<n_rows(matrix); i++) {
-//            transposed[j][i] = matrix[i][j];
-//        }
-//    }
-//    return transposed;
-//}
-//
+void transpose(const Matrix& matrix, Matrix& transposed) {
+    for (int j=0; j<n_cols(matrix); j++) {
+        for (int i=0; i<n_rows(matrix); i++) {
+            transposed[j][i] = matrix[i][j];
+        }
+    }
+}
+
 //// element-wise log
 //template <class T>
 //vector<vector<T>> log(const vector<vector<T>>& matrix) {

### 2017-02-07 matrix.cpp
Update log

@@ -66,7 +66,7 @@ void col_wise_sums(const Matrix& matrix, Matrix& result) {
     result.clear();
     for (int r = 0; r < matrix.n_rows; ++r) {
         for (int c = 0; c < matrix.n_cols; ++c)
-            result[c] += matrix[r][c];
+            result.data[0][c] += matrix.data[r][c];
     }
 }
 
@@ -87,22 +87,22 @@ void col_wise_sums(const Matrix& matrix, Matrix& result) {
 //
 //// matrix transposition
 void transpose(const Matrix& matrix, Matrix& transposed) {
-    for (int j=0; j<n_cols(matrix); j++) {
-        for (int i=0; i<n_rows(matrix); i++) {
-            transposed[j][i] = matrix[i][j];
+    for (int j=0; j < matrix.n_cols; j++) {
+        for (int i=0; i < matrix.n_rows; i++) {
+            transposed.data[j][i] = matrix.data[i][j];
         }
     }
 }
 
-//// element-wise log
-//template <class T>
-//vector<vector<T>> log(const vector<vector<T>>& matrix) {
-//    auto result = matrix;
-//    for (auto& row : result)
-//        for (auto& elem : row)
-//            elem = log(elem);
-//    return result;
-//}
+
+// element-wise log
+void log(const Matrix& matrix, Matrix& result) {
+    for (int i = 0; i < matrix.n_rows; ++i) {
+        for (int j = 0; j < matrix.n_cols; ++j) {
+            result.data[i][j] = log(matrix.data[i][j]);
+        }
+    }
+}
 //
 //// whole matrix sum
 //template <class T>

### 2017-02-07 matrix.cpp
Update sum-matrix

@@ -49,6 +49,8 @@ istream& operator>>(istream& is, Matrix& matrix) {
 }
 
 /*** Operators ***/
+// Adds first element of `to_add` to each element on the first column of `matrix`
+// second element of `to_add` to each of second column in `matrix`... etc
 void add_to_each(Matrix& matrix, const Matrix& to_add) {
     if (matrix.n_cols != to_add.n_cols)
         throw runtime_error(string_format("Add to each: matrix width and vector size are different: "
@@ -69,16 +71,6 @@ void col_wise_sums(const Matrix& matrix, Matrix& result) {
             result.data[0][c] += matrix.data[r][c];
     }
 }
-
-
-//Matrix chunk(const vector<T>& m, int from, int to) {
-//    if (from < 0)
-//        throw "From < 0";
-//    if (to > m.size()) to = m.size();
-//    return vector<T>(&m[from], &m[to]);
-//}
-
-
 
 //
 //
@@ -103,36 +95,17 @@ void log(const Matrix& matrix, Matrix& result) {
         }
     }
 }
-//
-//// whole matrix sum
-//template <class T>
-//T sum(const vector<vector<T>>& matrix) {
-//    T result = 0;
-//    for (const auto& row : matrix)
-//        for (const auto& elem : row)
-//            result += elem;
-//    return result;
-//}
-//
-//// Adds first element of `to_add` to each element on the first column of `matrix`
-//// second element of `to_add` to each of second column in `matrix`... etc
-//template <class T>
-//vector<vector<T>> add_to_each(const vector<vector<T>>& matrix, const vector<vector<T>>& to_add) {
-//    if (n_cols(matrix) != n_cols(to_add))
-//        throw runtime_error(string_format("Add to each: matrix width and vector size are different: "
-//                                                  "%d, %d", n_rows(matrix), n_rows(to_add)));
-//    if (n_rows(to_add) != 1)
-//        throw runtime_error(string_format("Add to each: vector to add has multiple rows: "
-//                                                  "%d", n_cols(to_add)));
-//
-//    vector<vector<T>> result = matrix;
-//    const vector<T>& v = to_add[0];
-//    for (int r = 0; r < n_rows(matrix); ++r)
-//        for (int c = 0; c < n_cols(matrix); ++c)
-//            result[r][c] += v[c];
-//
-//    return result;
-//}
+
+// whole matrix sum
[…]

### 2017-02-07 main.cpp
Resolve matrix errors

@@ -32,22 +32,32 @@ int predict(string pixels_string) {
 }
 
 int main(int argc, const char* argv[]) {
-    if (argc < 2) {
-        cout << "usage: mnist_in_mpi.out (train|predict)" << endl;
-        return 1;
-    }
+//    if (argc < 2) {
+//        cout << "usage: mnist_in_mpi.out (train|predict)" << endl;
+//        return 1;
+//    }
+//
+//    if (string(argv[1]) == "train") {
+//        train_and_save();
+//        return 0;
+//    }
+//
+//    if (string(argv[1]) == "predict") {
+//        if (argc < 3) {
+//            cout << "usage mnist_in_mpi.out predict <pixels>" << endl;
+//            return 1;
+//        }
+//        predict(argv[2]);
+//        return 0;
+//    }
 
-    if (string(argv[1]) == "train") {
-        train_and_save();
-        return 0;
-    }
 
-    if (string(argv[1]) == "predict") {
-        if (argc < 3) {
-            cout << "usage mnist_in_mpi.out predict <pixels>" << endl;
-            return 1;
-        }
-        predict(argv[2]);
-        return 0;
-    }
+    Matrix mat1(3, 4);
+    Matrix mat2(3, 4);
+
+
+
+
+
+    return 0;
 }

### 2017-02-07 matrix.cpp

@@ -23,7 +23,7 @@ public:
         delete[] data;
     }
 
-    clear() {
+    void clear() {
         for (int i = 0; i < n_rows; ++i)
             for (int j = 0; j < n_cols; ++j)
                 data[i][j] = 0;
@@ -36,7 +36,7 @@ public:
 ostream& operator<<(ostream& os, const Matrix& matrix) {
     for (int i = 0; i < matrix.n_rows; ++i) {
         for (int j = 0; j < matrix.n_cols; ++j)
-            os << matrix[i][j] << ' ';
+            os << matrix.data[i][j] << ' ';
         os << endl;
     }
     return os;
@@ -44,7 +44,7 @@ ostream& operator<<(ostream& os, const Matrix& matrix) {
 istream& operator>>(istream& is, Matrix& matrix) {
     for (int i = 0; i < matrix.n_rows; ++i)
         for (int j = 0; j < matrix.n_cols; ++j)
-            is >> matrix[i][j];
+            is >> matrix.data[i][j];
     return is;
 }
 
@@ -54,10 +54,10 @@ istream& operator>>(istream& is, Matrix& matrix) {
 void add_to_each(Matrix& matrix, const Matrix& to_add) {
     if (matrix.n_cols != to_add.n_cols)
         throw runtime_error(string_format("Add to each: matrix width and vector size are different: "
-                                                  "%d, %d", n_rows(matrix), n_rows(to_add)));
+                                                  "%d, %d", matrix.n_rows, matrix.n_rows));
     if (to_add.n_rows != 1)
         throw runtime_error(string_format("Add to each: vector to add has multiple rows: "
-                                                  "%d", n_cols(to_add)));
+                                                  "%d", to_add.n_rows));
 
     for (int r = 0; r < to_add.n_rows; ++r)
         for (int c = 0; c < matrix.n_cols; ++c)
@@ -98,7 +98,7 @@ void log(const Matrix& matrix, Matrix& result) {
 
 // whole matrix sum
 void sum(const Matrix& matrix, double result) {
-    double result = 0;
+    result = 0;
     for (int i = 0; i < matrix.n_rows ; ++i) {
         for (int j = 0; j < matrix.n_cols; ++j) {
             result += matrix.data[i][j];
@@ -107,15 +107,10 @@ void sum(const Matrix& matrix, double result) {
 }
 
 //
-//
-//
-//
-//
 ///*** Operators ***/
 //
[…]

### 2017-02-07 main.cpp
Reading

@@ -1,63 +1,0 @@
-#include <iostream>
-#include "reading.cpp"
-#include "nn.cpp"
-
-using namespace std;
-
-void train_and_save() {
-    // TODO: names as additional parameters
-    Matrix images, labels;
-    read_data("data/train.csv", images, labels);
-
-    NeuralNetwork net;
-    vector<double> costs, accs;
-    net.train(images, labels, costs, accs, 35, 100, .1);
-
-    net.save("models/trained.nn");
-    ofstream f("histories/training.txt");
-    f << costs << endl << accs;
-    f.close();
-}
-
-int predict(string pixels_string) {
-    // TODO: names as additional parameters
-    vector<double> pixels = pixels_from_string(pixels_string);
-    NeuralNetwork net("models/trained_a.nn");
-
-    double confidence;
-    int digit_predicted = net.predict_one(pixels, confidence);
-    cout << digit_predicted << endl << confidence << endl;
-
-    return digit_predicted;
-}
-
-int main(int argc, const char* argv[]) {
-//    if (argc < 2) {
-//        cout << "usage: mnist_in_mpi.out (train|predict)" << endl;
-//        return 1;
-//    }
-//
-//    if (string(argv[1]) == "train") {
-//        train_and_save();
-//        return 0;
-//    }
-//
-//    if (string(argv[1]) == "predict") {
-//        if (argc < 3) {
-//            cout << "usage mnist_in_mpi.out predict <pixels>" << endl;
-//            return 1;
-//        }
-//        predict(argv[2]);
-//        return 0;
-//    }
-
-
-    Matrix mat1(3, 4);
-    Matrix mat2(3, 4);
-
-
-
-
-
-    return 0;
-}

### 2017-02-07 nn.cpp

@@ -63,6 +63,14 @@ class NeuralNetwork {
     Matrix W;
     Matrix b;
 
+    Matrix Y_prob;
+    Matrix delta;
+    Matrix grad_W;
+    Matrix grad_b;
+    Matirx XW;
+    Matrix XWb;
+    Matrix softmaxed;
+
 public:
 
     NeuralNetwork(int n_classes=10, int data_dim=784)
@@ -80,6 +88,9 @@ public:
     double grad(Matrix X, Matrix Y,
                 Matrix &grad_W, Matrix &grad_b) {
         double n = n_rows(X);  // double instead of int so division works
+        mult(X, W, XW);
+        add(XW, b, XWb);
+        softmax(XWb, softmaxed);
 
         Matrix Y_prob = softmax(X * W + b); // add bias to each
         // Error at last layer

### 2017-02-07 reading.cpp

@@ -10,50 +10,50 @@
 using namespace std;
 
 const int PIXEL_DIM = 255;
+const int IMAGE_SIDE = 28;
 
-void read_data(string filename, Matrix& images, Matrix& labels) {
+void read_from_csv(string filename, Matrix& images, Matrix& labels) {
+    // images.n_rows indicates how many images to read from the csv
     ifstream file(filename);
 
     if (!file.is_open())
         throw runtime_error(string_format("Read MNIST data: could not open file: " + filename));
 
-    vector<int> label_values;
+    images.clear();
+    labels.clear();
+
+    int label_value;
     string line;
-    while (getline(file, line)) {
+    for (int image_n = 0; image_n < images.n_rows; ++image_n) {
+        getline(file, line);
         istringstream line_stream(line);
 
-        int label;
-        line_stream >> label;
-        label_values.push_back(label);
+        // Read label
+        line_stream >> label_value;
+        labels.data[image_n][label_value] = 1; // one-hot encoding
         line_stream.ignore(); // skip comma after label
 
-        vector<double> image;
-        int pixel;
-        while (line_stream >> pixel) {
-            image.push_back(double(pixel) / PIXEL_DIM);
+        // Read each pixel
+        for (int pixel_n = 0; pixel_n < images.n_cols; ++pixel_n) {
+            line_stream >> images.data[image_n][pixel_n];
             line_stream.ignore(); // skip comma after pixel value
         }
-
-        images.push_back(image);
     }
     file.close();
-
-    for (auto val : label_values)
-        labels.push_back(make_one_hot(val));
 
     cout << "Done reading from " + filename << endl;
 }
 
-void print_image(const vector<double>& pixels) {
-    for (int i = 0; i < 28 * 28; ++i) {
-        if (i > 0 && i % 28 == 0)
-            cout << endl << endl;
+void print_image(const Matrix& images, const int image_n) {
+    for (int pixel_n = 0; pixel_n < IMAGE_SIDE * IMAGE_SIDE; ++pixel_n) {
+        if (i % IMAGE_SIDE == 0)
+            cout << endl;
 
-        if (pixels[i] == 0)
+        if (images.data[image_n][pixel_n] == 0)
             cout << "  ";
         else
-            cout << int(pixels[i] * 100); // turns 0.25 into 25
+            cout << int(images.data[image_n][pixel_n] * 100);
         cout << ' ';
     }
-    cout << endl << endl;
+    cout << endl;
 }

### 2017-02-07 utils.cpp

@@ -28,37 +28,37 @@ std::string string_format(const std::string fmt_str, ...) {
     return std::string(formatted.get());
 }
 
-template <class T>
-ostream& operator<<(ostream& os, const vector<T>& v) {
-    for (const auto& x : v)
-        os << x << ' ';
-    return os;
-}
-
-template <class T>
-istream& operator>>(istream& is, vector<T>& v) {
-    string line;
-    getline(is, line);
-    istringstream ss(line);
-
-    T x;
-    while (ss >> x) // until end of line
-        v.push_back(x);
-
-    return is;
-}
-
-vector<double> pixels_from_string(const string& s) {
-    vector<double> v;
-    istringstream ss(s);
-    double x;
-    while (ss >> x) {
-        v.push_back(x);
-        if (ss.peek() == ',')
-            ss.ignore();
-    }
-    return v;
-}
+//template <class T>
+//ostream& operator<<(ostream& os, const vector<T>& v) {
+//    for (const auto& x : v)
+//        os << x << ' ';
+//    return os;
+//}
+//
+//template <class T>
+//istream& operator>>(istream& is, vector<T>& v) {
+//    string line;
+//    getline(is, line);
+//    istringstream ss(line);
+//
+//    T x;
+//    while (ss >> x) // until end of line
+//        v.push_back(x);
+//
+//    return is;
+//}
+//
+//vector<double> pixels_from_string(const string& s) {
+//    vector<double> v;
+//    istringstream ss(s);
+//    double x;
+//    while (ss >> x) {
+//        v.push_back(x);
+//        if (ss.peek() == ',')
+//            ss.ignore();
+//    }
+//    return v;
+//}
 
 
 // for 2 (and 5 classes) ~> {0 0 1 0 0}

### 2017-02-07 reading.cpp
Fix reading

@@ -1,9 +1,9 @@
 #pragma once
 
-#include <iostream>
 #include <string>
 #include <fstream>
 #include <sstream>
+#include <iostream>
 #include "matrix.cpp"
 #include "utils.cpp"
 
@@ -36,6 +36,7 @@ void read_from_csv(string filename, Matrix& images, Matrix& labels) {
         // Read each pixel
         for (int pixel_n = 0; pixel_n < images.n_cols; ++pixel_n) {
             line_stream >> images.data[image_n][pixel_n];
+            images.data[image_n][pixel_n] /= PIXEL_DIM;
             line_stream.ignore(); // skip comma after pixel value
         }
     }
@@ -46,8 +47,8 @@ void read_from_csv(string filename, Matrix& images, Matrix& labels) {
 
 void print_image(const Matrix& images, const int image_n) {
     for (int pixel_n = 0; pixel_n < IMAGE_SIDE * IMAGE_SIDE; ++pixel_n) {
-        if (i % IMAGE_SIDE == 0)
-            cout << endl;
+        if (pixel_n % IMAGE_SIDE == 0)
+            cout << endl << endl;
 
         if (images.data[image_n][pixel_n] == 0)
             cout << "  ";

### 2017-02-07 utils.cpp

@@ -62,11 +62,11 @@ std::string string_format(const std::string fmt_str, ...) {
 
 
 // for 2 (and 5 classes) ~> {0 0 1 0 0}
-vector<double> make_one_hot(int value, int n_classes=10) {
-    vector<double> result(n_classes, 0);
-    result[value] = 1;
-    return result;
-}
+//vector<double> make_one_hot(int value, int n_classes=10) {
+//    vector<double> result(n_classes, 0);
+//    result[value] = 1;
+//    return result;
+//}
 
 // {0 0 1 0 0} ~> 2
 int from_one_hot(const vector<double>& v) {

### 2017-02-07 nn.cpp
nn sketch

@@ -67,9 +67,18 @@ class NeuralNetwork {
     Matrix delta;
     Matrix grad_W;
     Matrix grad_b;
+
     Matirx XW;
     Matrix XWb;
-    Matrix softmaxed;
+    Matrix Y_prob;
+
+    Matrix delta;
+
+    Matrix trX;
+    Matrix trXdelta;
+    Matrix grad_W;
+
+    Matrix delta_sums;
 
 public:
 
@@ -84,23 +93,24 @@ public:
         this->load(path);
     }
 
-    // TODO: parallelization
-    double grad(Matrix X, Matrix Y,
-                Matrix &grad_W, Matrix &grad_b) {
-        double n = n_rows(X);  // double instead of int so division works
-        mult(X, W, XW);
-        add(XW, b, XWb);
-        softmax(XWb, softmaxed);
+    void grad(Matrix X, Matrix Y, Matrix& grad_W, Matrix& grad_b) {
+        dot(X, W, XW); // XW = X * W
+        add_to_each(XW, b, XWb); // XWb = X * W + b
+        softmax(XWb, Y_prob); // softmaxed = softmax(X * W + b)
+        sub(Y_prob, Y, delta); // delta = Y_prob - Y
 
-        Matrix Y_prob = softmax(X * W + b); // add bias to each
-        // Error at last layer
-        Matrix delta = Y_prob - Y;
+        double contribution = 1/double(X.n_rows);
 
-        // Gradient of cross entropy cost
-        grad_W = 1/n * transpose(X) * delta;
-        grad_b = 1/n * col_wise_sums(delta);
+        transpose(X, trX); // trX = transpose(X)
+        dot(trX, delta, trXdelta); // trXdelta = transpose(X) * delta
+        scalar_mult(contribution, trXdelta, grad_W); // grad_W = 1/n * transpose(X) * delta
 
-        return cross_entropy(Y, Y_prob); // cost
+
+        col_wise_sums(delta, delta_sums);
+        scalar_mult(contribution, delta_sums); // grad_b = 1/n * col_wise_sums(delta)
+
+        // TODO maybe
+//        return cross_entropy(Y, Y_prob); // cost
     }
 
     void train(Matrix X, Matrix Y,

### 2017-02-07 main.cpp
Add data_anchor

@@ -1,0 +1,72 @@
+#include <iostream>
+#include "reading.cpp"
+//#include "nn.cpp"
+#include "matrix.cpp"
+
+using namespace std;
+
+//void train_and_save() {
+//    // TODO: names as additional parameters
+//    Matrix images, labels;
+//    read_data("data/train.csv", images, labels);
+//
+//    NeuralNetwork net;
+//    vector<double> costs, accs;
+//    net.train(images, labels, costs, accs, 35, 100, .1);
+//
+//    net.save("models/trained.nn");
+//    ofstream f("histories/training.txt");
+//    f << costs << endl << accs;
+//    f.close();
+//}
+
+//int predict(string pixels_string) {
+//    // TODO: names as additional parameters
+//    vector<double> pixels = pixels_from_string(pixels_string);
+//    NeuralNetwork net("models/trained_a.nn");
+//
+//    double confidence;
+//    int digit_predicted = net.predict_one(pixels, confidence);
+//    cout << digit_predicted << endl << confidence << endl;
+//
+//    return digit_predicted;
+//}
+
+int main(int argc, const char* argv[]) {
+//    if (argc < 2) {
+//        cout << "usage: mnist_in_mpi.out (train|predict)" << endl;
+//        return 1;
+//    }
+//
+//    if (string(argv[1]) == "train") {
+//        train_and_save();
+//        return 0;
+//    }
+//
+//    if (string(argv[1]) == "predict") {
+//        if (argc < 3) {
+//            cout << "usage mnist_in_mpi.out predict <pixels>" << endl;
+//            return 1;
+//        }
+//        predict(argv[2]);
+//        return 0;
+//    }
+
+    int n_samples = 250;
+    Matrix images(n_samples, 784), labels(n_samples, 10);
+    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/sample.csv"), images, labels);
+    print_image(images, 0);
+
+    Matrix mat1(3, 4);
+    Matrix mat2(3, 4);
+
+    Matrix res(3, 4);
+
+    add(mat1, mat2, res);
+
+    cout << res;
+
+
+
+    return 0;
+}

### 2017-02-07 matrix.cpp

@@ -2,11 +2,12 @@
 
 #include <math.h>  // exp, math
 #include "utils.cpp"
-
+#include <iostream>
 
 class Matrix {
 
 public:
+    double** data_anchor;
     double** data;
     int n_rows;
     int n_cols;
@@ -15,12 +16,18 @@ public:
         data = new double*[n_rows];
         for (int i = 0; i < n_rows; ++i)
             data[i] = new double[n_cols];
+        data_anchor = data;  // Set checkpoint to start of data
     }
 
     ~Matrix() {
+        this->reset_to_anchor();
         for (int i = 0; i < n_rows; ++i)
             delete[] data[i];
         delete[] data;
+    }
+
+    void reset_to_anchor() {
+        this->data = data_anchor;
     }
 
     void clear() {
@@ -33,15 +40,15 @@ public:
 
 
 /*** IO ***/
-ostream& operator<<(ostream& os, const Matrix& matrix) {
+std::ostream& operator<<(std::ostream& os, const Matrix& matrix) {
     for (int i = 0; i < matrix.n_rows; ++i) {
         for (int j = 0; j < matrix.n_cols; ++j)
             os << matrix.data[i][j] << ' ';
-        os << endl;
+        os << std::endl;
     }
     return os;
 }
-istream& operator>>(istream& is, Matrix& matrix) {
+std::istream& operator>>(std::istream& is, Matrix& matrix) {
     for (int i = 0; i < matrix.n_rows; ++i)
         for (int j = 0; j < matrix.n_cols; ++j)
             is >> matrix.data[i][j];
@@ -51,7 +58,7 @@ istream& operator>>(istream& is, Matrix& matrix) {
 /*** Operators ***/
 // Adds first element of `to_add` to each element on the first column of `matrix`
 // second element of `to_add` to each of second column in `matrix`... etc
-void add_to_each(Matrix& matrix, const Matrix& to_add) {
+void add_to_each(Matrix& matrix, const Matrix& to_add, Matrix& result) {
     if (matrix.n_cols != to_add.n_cols)
         throw runtime_error(string_format("Add to each: matrix width and vector size are different: "
                                                   "%d, %d", matrix.n_rows, matrix.n_rows));
@@ -61,7 +68,7 @@ void add_to_each(Matrix& matrix, const Matrix& to_add) {
 
[…]

### 2017-02-07 main.cpp
Matrix chunk

@@ -52,20 +52,21 @@ int main(int argc, const char* argv[]) {
 //        return 0;
 //    }
 
-    int n_samples = 250;
-    Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/sample.csv"), images, labels);
-    print_image(images, 0);
+//    int n_samples = 250;
+//    Matrix images(n_samples, 784), labels(n_samples, 10);
+//    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/sample.csv"), images, labels);
+//    print_image(images, 0);
 
-    Matrix mat1(3, 4);
-    Matrix mat2(3, 4);
-
-    Matrix res(3, 4);
-
-    add(mat1, mat2, res);
-
-    cout << res;
-
+//    Matrix m(4, 2);
+//    int k = 0;
+//    for (int i = 0; i < m.n_rows; ++i)
+//        for (int j = 0; j < m.n_cols; ++j)
+//            m.data[i][j] = k++;
+//    cout << m;
+//
+//    Matrix chunk(2, 2);
+//    take_chunk(m, 1, chunk);
+//    cout << endl << chunk;
 
 
     return 0;

### 2017-02-07 matrix.cpp

@@ -7,27 +7,23 @@
 class Matrix {
 
 public:
-    double** data_anchor;
     double** data;
     int n_rows;
     int n_cols;
+    bool is_chunk = false;
 
     Matrix(int n_rows,int n_cols): n_rows(n_rows), n_cols(n_cols) {
         data = new double*[n_rows];
         for (int i = 0; i < n_rows; ++i)
             data[i] = new double[n_cols];
-        data_anchor = data;  // Set checkpoint to start of data
     }
 
     ~Matrix() {
-        this->reset_to_anchor();
+        if (is_chunk)
+            return;
         for (int i = 0; i < n_rows; ++i)
             delete[] data[i];
         delete[] data;
-    }
-
-    void reset_to_anchor() {
-        this->data = data_anchor;
     }
 
     void clear() {
@@ -79,12 +75,13 @@ void col_wise_sums(const Matrix& matrix, Matrix& result) {
     }
 }
 
-//
-//
-//
-///*** Operations ***/
-//
-//// matrix transposition
+/*** Operations ***/
+inline void take_chunk(const Matrix& from, int start_index, Matrix& into) {
+    into.data = from.data + start_index * from.n_cols; // skip n rows, where n is the start index
+    into.is_chunk = true;
+}
+
+// matrix transposition
 void transpose(const Matrix& matrix, Matrix& transposed) {
     for (int j=0; j < matrix.n_cols; j++) {
         for (int i=0; i < matrix.n_rows; i++) {

### 2017-02-07 nn.cpp

@@ -80,6 +80,9 @@ class NeuralNetwork {
 
     Matrix delta_sums;
 
+    Matrix chunk_X(chunk_size, 784);
+
+
 public:
 
     NeuralNetwork(int n_classes=10, int data_dim=784)
@@ -120,15 +123,17 @@ public:
         double cost, acc;
         auto Y_labels = from_one_hot_matrix(Y); // {5, 2, 9, ... }
 
-        Matrix grad_W = blank_matrix(data_dim, n_classes, 0.);
-        Matrix grad_b = blank_matrix(1, n_classes, 0.);
+//        Matrix grad_W = blank_matrix(data_dim, n_classes, 0.);
+//        Matrix grad_b = blank_matrix(1, n_classes, 0.);
 
         for (int epoch = 1; epoch <= n_epochs; ++epoch) {
             if (verbose)
                 cout << string_format("Epoch %d / %d", epoch, n_epochs) << endl;
 
+
             for (int i = 0; i < n_rows(X); i+=batch_size) {
                 // TODO: make random batch generator
+
                 cost = grad(chunk(X, i, i+batch_size), chunk(Y, i, i+batch_size),
                             grad_W,  grad_b);
 

### 2017-02-07 nn.cpp
nn train sketch

@@ -58,10 +58,14 @@ Matrix random_init(int rows, int cols, double mean=0., double std=1.) {
 
 class NeuralNetwork {
 
-    int n_classes;
-    int data_dim;
-    Matrix W;
-    Matrix b;
+    const int N_CLASSES = 10; // 10 digits from 0 to 9
+    const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
+    const int BATCH_SIZE = 100; // after how many pictures the weights are updated
+    const int N_WORKERS = 4; // how many processes compute gradient in parallel
+    const int CHUNK_SIZE = BATCH_SIZE / N_WORKERS; // how many rows each worker gets
+
+    Matrix W = Matrix(DATA_DIM, N_CLASSES);
+    Matrix b = Matrix(1, N_CLASSES);
 
     Matrix Y_prob;
     Matrix delta;
@@ -80,7 +84,11 @@ class NeuralNetwork {
 
     Matrix delta_sums;
 
-    Matrix chunk_X(chunk_size, 784);
+    Matrix chunk_X = Matrix(CHUNK_SIZE, DATA_DIM);
+    Matrix chunk_Y = Matrix(CHUNK_SIZE, N_CLASSES);
+
+    Matrix lr_grad_W;
+    matrix lr_grad_b;
 
 
 public:
@@ -96,10 +104,10 @@ public:
         this->load(path);
     }
 
-    void grad(Matrix X, Matrix Y, Matrix& grad_W, Matrix& grad_b) {
+    void grad(const Matrix& X, const Matrix& Y) {
         dot(X, W, XW); // XW = X * W
         add_to_each(XW, b, XWb); // XWb = X * W + b
-        softmax(XWb, Y_prob); // softmaxed = softmax(X * W + b)
+        softmax(XWb, Y_prob); // Y_prob = softmax(X * W + b)
         sub(Y_prob, Y, delta); // delta = Y_prob - Y
 
         double contribution = 1/double(X.n_rows);
@@ -110,18 +118,18 @@ public:
 
 
         col_wise_sums(delta, delta_sums);
-        scalar_mult(contribution, delta_sums); // grad_b = 1/n * col_wise_sums(delta)
+        scalar_mult(contribution, delta_sums, grad_b); // grad_b = 1/n * col_wise_sums(delta)
 
         // TODO maybe
 //        return cross_entropy(Y, Y_prob); // cost
     }
 
-    void train(Matrix X, Matrix Y,
+    void train(const Matrix& X, const Matrix& Y,
[…]

### 2017-02-07 matrix.cpp
Add dot

@@ -128,20 +128,19 @@ void add (const Matrix& lhs, const Matrix& rhs, Matrix& result) {
         for (int j = 0; j < rhs.n_cols; ++j)
             result.data[i][j] = rhs.data[i][j] + lhs.data[i][j];
 }
-//
-//// matrix - vector addition
-//template <class T>
-//vector<vector<T>> operator+ (const vector<vector<T>>& matrix, const vector<T>& vect) {
-//    if (n_cols(matrix) != vect.size())
-//        throw runtime_error(string_format("Matrix - vector addition: number of rows is different: "
-//                                                  "%d, %d", n_cols(matrix), vect.size()));
-//
-//    auto result = matrix;
-//    for (int i = 0; i < n_rows(matrix); ++i)
-//        for (int j = 0; j < n_cols(matrix); ++j)
-//            result[i][j] += vect[j];
-//    return result;
-//}
+
+// matrix - vector addition
+void add_vect (const Matrix& matrix, const Matrix& vect, Matrix& result) {
+    if (matrix.n_cols != vect.n_cols)
+        throw runtime_error(string_format("Matrix - vector addition: number of rows is different: "
+                                                  "%d, %d", matrix.n_cols, vect.n_cols));
+    if (vect.n_rows != 1)
+        throw runtime_error(string_format("Matrix - vector addition: vector dims: "
+                                                  "%d, %d", vect.n_rows, vect.n_cols));
+    for (int i = 0; i < matrix.n_rows; ++i)
+        for (int j = 0; j < matrix.n_cols; ++j)
+            result.data[i][j] = matrix.data[i][j] + vect.data[0][j];
+}
 //
 //
 //// matrix subtraction
@@ -171,25 +170,21 @@ void add (const Matrix& lhs, const Matrix& rhs, Matrix& result) {
 //
 //// TODO: faster method
 //// matrix multiplication
-//template <class T>
-//vector<vector<T>> operator* (const vector<vector<T>>& lhs, const vector<vector<T>>& rhs) {
-//    if (n_cols(lhs) != n_rows(rhs))
-//        throw runtime_error(string_format("Dimensions do not agree for matrix multiplication: "
-//                      "lhs cols = %d, rhs rows = %d", n_cols(lhs), n_rows(rhs)));
-//
-//    int n = n_rows(lhs);
-//    int p = n_cols(lhs);
[…]

### 2017-02-07 matrix.cpp
sub and sub_from

@@ -114,9 +114,9 @@ void sum(const Matrix& matrix, double result) {
 ///*** Operators ***/
 //
 //// matrix addition
-void add (const Matrix& lhs, const Matrix& rhs, Matrix& result) {
+void add(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
     // If rhs is a row-vector, add it to each row of the lhs
-    if (rhs.n_rows == 1 && rhs.n_cols == lhs.n_cols)
+//    if (rhs.n_rows == 1 && rhs.n_cols == lhs.n_cols)
 //        add_to_each(lhs, rhs, result);
 
     if ((lhs.n_rows != rhs.n_rows) || (lhs.n_cols != rhs.n_cols))
@@ -128,6 +128,29 @@ void add (const Matrix& lhs, const Matrix& rhs, Matrix& result) {
         for (int j = 0; j < rhs.n_cols; ++j)
             result.data[i][j] = rhs.data[i][j] + lhs.data[i][j];
 }
+
+void sub(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
+    if ((lhs.n_rows != rhs.n_rows) || (lhs.n_cols != rhs.n_cols))
+        throw runtime_error(string_format("Matrix subtraction: number of rows/cols is different: "
+                                                  "lhs = (%d, %d), rhs = (%d, %d)",
+                                          lhs.n_rows, lhs.n_cols, rhs.n_rows, rhs.n_cols));
+
+    for (int i = 0; i < rhs.n_rows; ++i)
+        for (int j = 0; j < rhs.n_cols; ++j)
+            result.data[i][j] = rhs.data[i][j] - lhs.data[i][j];
+}
+
+void sub(Matrix& from, const Matrix& to_sub) {
+    if ((from.n_rows != to_sub.n_rows) || (from.n_cols != to_sub.n_cols))
+        throw runtime_error(string_format("Matrix FROM subtraction: number of rows/cols is different: "
+                                                  "from = (%d, %d), to_sub = (%d, %d)",
+                                          from.n_rows, from.n_cols, to_sub.n_rows, to_sub.n_cols));
+
+    for (int i = 0; i < from.n_rows; ++i)
+        for (int j = 0; j < from.n_cols; ++j)
+            from.data[i][j] -= to_sub.data[i][j];
+}
+
 
 // matrix - vector addition
 void add_vect (const Matrix& matrix, const Matrix& vect, Matrix& result) {
[…]

### 2017-02-07 matrix.cpp
amend

@@ -140,7 +140,8 @@ void sub(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
             result.data[i][j] = rhs.data[i][j] - lhs.data[i][j];
 }
 
-void sub(Matrix& from, const Matrix& to_sub) {
+void sub_from(Matrix& from, const Matrix& to_sub) {
+    // from = from - to_sub
     if ((from.n_rows != to_sub.n_rows) || (from.n_cols != to_sub.n_cols))
         throw runtime_error(string_format("Matrix FROM subtraction: number of rows/cols is different: "
                                                   "from = (%d, %d), to_sub = (%d, %d)",

### 2017-02-07 matrix.cpp
Add hadamard

@@ -34,7 +34,6 @@ public:
 };
 
 
-
 /*** IO ***/
 std::ostream& operator<<(std::ostream& os, const Matrix& matrix) {
     for (int i = 0; i < matrix.n_rows; ++i) {
@@ -115,10 +114,6 @@ void sum(const Matrix& matrix, double result) {
 //
 //// matrix addition
 void add(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
-    // If rhs is a row-vector, add it to each row of the lhs
-//    if (rhs.n_rows == 1 && rhs.n_cols == lhs.n_cols)
-//        add_to_each(lhs, rhs, result);
-
     if ((lhs.n_rows != rhs.n_rows) || (lhs.n_cols != rhs.n_cols))
         throw runtime_error(string_format("Matrix addition: number of rows/cols is different: "
                                                   "lhs = (%d, %d), rhs = (%d, %d)",
@@ -165,19 +160,17 @@ void add_vect (const Matrix& matrix, const Matrix& vect, Matrix& result) {
         for (int j = 0; j < matrix.n_cols; ++j)
             result.data[i][j] = matrix.data[i][j] + vect.data[0][j];
 }
-//
-//// matrix-scalar multiplication
-//template <class T>
-//vector<vector<T>> operator* (double scalar, const vector<vector<T>>& matrix) {
-//    auto result = matrix;
-//    for (auto& row : result)
-//        for (auto& elem : row)
-//            elem *= scalar;
-//    return result;
-//}
-//
-//// TODO: faster method
-//// matrix multiplication
+
+// matrix-scalar multiplication
+void scalar_mult(double scalar, const Matrix& matrix, Matrix& result) {
+    for (int i = 0; i < matrix.n_rows; ++i) {
+        for (int j = 0; j < matrix.n_cols; ++j) {
+            result.data[i][j] = scalar * matrix.data[i][j];
+        }
+    }
+}
+
+// matrix multiplication
 void dot(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
     if (lhs.n_cols != rhs.n_rows)
         throw runtime_error(string_format("Dimensions do not agree for matrix multiplication: "
@@ -193,17 +186,14 @@ void dot(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
                 result.data[i][j] += lhs.data[i][k] * rhs.data[k][j];
 }
 
-//// element-wise matrix multiplication
-//template <class T>
[…]

### 2017-02-07 nn.cpp
softmax & random init

@@ -11,46 +11,38 @@ using namespace std;
 
 
 /*** Requisites ***/
-vector<double> softmax(const vector<double>& v) {
-    // Subtract maximum to avoid overflow
-    double max = *max_element(v.begin(), v.end());
+void softmax(const Matrix& matrix, Matrix& result) {
+    for (int row = 0; row < matrix.n_rows; ++row) {
+        double row_max = matrix[row][0];
+        for (int i = 1; i < n; ++i)
+            row_max = matrix[row][i] > row_max ? matrix[row][i] : row_max;
 
-    vector<double> expd = v; // copy, does not mutate v
-    double sum = 0;
-    for (double& x : expd) {
-        x = exp(x - max);
-        sum += x;
+        double row_sum = 0;
+        // Subtract maximum to avoid overflow
+        for (int i = 0; i < n; ++i) {
+            result[row][i] = exp(matrix[row][i] - row_max);
+            row_sum += result[row][i];
+        }
+
+        for (int i = 0; i < n; ++i)
+            result[row][i] /= row_sum;
     }
-
-    vector<double> result(v.size(), 0.);
-    for (int i = 0; i < v.size(); ++i)
-        result[i] = expd[i] / sum;
-
-    return result;
-}
-Matrix softmax(const Matrix& m) {
-    Matrix result = m;
-    for (auto& row : result)
-        row = softmax(row);
-    return result;
 }
 
-double cross_entropy(const Matrix& Y, const Matrix& Y_prob) {
-    // Cost function
-    Matrix result = hadamard(Y, log(Y_prob));
-    return 1/double(n_rows(Y)) * sum(result);
-}
+// TODO
+//double cross_entropy(const Matrix& Y, const Matrix& Y_prob) {
+//    // Cost function
+//    Matrix result = hadamard(Y, log(Y_prob));
+//    return 1/double(n_rows(Y)) * sum(result);
+//}
 
-Matrix random_init(int rows, int cols, double mean=0., double std=1.) {
+void random_init(Matrix& W, double mean=0., double std=1.) {
     default_random_engine generator;
     normal_distribution<double> distribution(mean, std);
 
-    Matrix result = blank_matrix(rows, cols, 0.);
-    for (auto& row : result)
-        for (double& elem : row)
-            elem = distribution(generator);
-
-    return result;
+    for (int i = 0; i < W.n_rows; ++i)
[…]

### 2017-02-07 main.cpp
Debug in progress

@@ -1,6 +1,6 @@
 #include <iostream>
 #include "reading.cpp"
-//#include "nn.cpp"
+#include "nn.cpp"
 #include "matrix.cpp"
 
 using namespace std;
@@ -52,10 +52,12 @@ int main(int argc, const char* argv[]) {
 //        return 0;
 //    }
 
-//    int n_samples = 250;
-//    Matrix images(n_samples, 784), labels(n_samples, 10);
-//    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/sample.csv"), images, labels);
-//    print_image(images, 0);
+    int n_samples = 250;
+    Matrix images(n_samples, 784), labels(n_samples, 10);
+    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/sample.csv"), images, labels);
+    NeuralNetwork net;
+    net.train(images, labels);
+
 
 //    Matrix m(4, 2);
 //    int k = 0;

### 2017-02-07 matrix.cpp

@@ -33,6 +33,9 @@ public:
     }
 };
 
+void print_dimensions(const Matrix& m) {
+    cout << m.n_rows << " x " << m.n_cols << endl;
+}
 
 /*** IO ***/
 std::ostream& operator<<(std::ostream& os, const Matrix& matrix) {

### 2017-02-07 nn.cpp

@@ -4,7 +4,7 @@
 #include <iostream>
 #include <fstream>
 #include "matrix.cpp"
-#include "evaluate.cpp"
+//#include "evaluate.cpp" // TODO
 
 
 using namespace std;
@@ -13,19 +13,19 @@ using namespace std;
 /*** Requisites ***/
 void softmax(const Matrix& matrix, Matrix& result) {
     for (int row = 0; row < matrix.n_rows; ++row) {
-        double row_max = matrix[row][0];
-        for (int i = 1; i < n; ++i)
-            row_max = matrix[row][i] > row_max ? matrix[row][i] : row_max;
+        double row_max = matrix.data[row][0];
+        for (int i = 1; i < matrix.n_cols; ++i)
+            row_max = matrix.data[row][i] > row_max ? matrix.data[row][i] : row_max;
 
         double row_sum = 0;
         // Subtract maximum to avoid overflow
-        for (int i = 0; i < n; ++i) {
-            result[row][i] = exp(matrix[row][i] - row_max);
-            row_sum += result[row][i];
+        for (int i = 0; i < matrix.n_cols; ++i) {
+            result.data[row][i] = exp(matrix.data[row][i] - row_max);
+            row_sum += result.data[row][i];
         }
 
-        for (int i = 0; i < n; ++i)
-            result[row][i] /= row_sum;
+        for (int i = 0; i < matrix.n_cols; ++i)
+            result.data[row][i] /= row_sum;
     }
 }
 
@@ -50,61 +50,67 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 
 class NeuralNetwork {
 
+    const int N_SAMPLES = 250;
     const int N_CLASSES = 10; // 10 digits from 0 to 9
     const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
     const int BATCH_SIZE = 100; // after how many pictures the weights are updated
     const int N_WORKERS = 4; // how many processes compute gradient in parallel
     const int CHUNK_SIZE = BATCH_SIZE / N_WORKERS; // how many rows each worker gets
 
-    Matrix W = Matrix(DATA_DIM, N_CLASSES);
+    // X dimensions: N_SAMPLES x DATA_DIM
+    // Y dimensions: N_SAMPLES x N_CLASSES
+
+    Matrix W = Matrix(DATA_DIM, N_CLASSES);        // TODO? more layers
     Matrix b = Matrix(1, N_CLASSES);
-
-    Matrix Y_prob;
-    Matrix delta;
[…]

### 2017-02-07 main.cpp
Fix take_chunk

@@ -52,6 +52,12 @@ int main(int argc, const char* argv[]) {
 //        return 0;
 //    }
 
+//    double* v = new double[5];
+//    for (int i = 0; i < 5; ++i)
+//        v[i] = 100+i;
+//    double* w = v+2;
+//    cout << w[0];
+
     int n_samples = 250;
     Matrix images(n_samples, 784), labels(n_samples, 10);
     read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/sample.csv"), images, labels);

### 2017-02-07 matrix.cpp

@@ -78,8 +78,8 @@ void col_wise_sums(const Matrix& matrix, Matrix& result) {
 }
 
 /*** Operations ***/
-inline void take_chunk(const Matrix& from, int start_index, Matrix& into) {
-    into.data = from.data + start_index * from.n_cols; // skip n rows, where n is the start index
+inline void take_chunk(const Matrix& from, int start_row, Matrix& into) {
+    into.data = &(from.data[start_row]);
     into.is_chunk = true;
 }
 

### 2017-02-07 nn.cpp

@@ -92,11 +92,19 @@ public:
 //    }
 
     void grad() {
-        print_dimensions(chunk_X);
-        print_dimensions(W);
-        print_dimensions(XW);
+//        print_dimensions(chunk_X);
+//        print_dimensions(W);
+//        print_dimensions(XW);
+//        cout << "first elem of chunk x = " << chunk_X.data[0][0] << endl;
+//        for (int i = 0; i < chunk_X.n_rows; ++i) {
+//            for (int j = 0; j < chunk_X.n_cols; ++j) {
+//                chunk_X.data[i][j] *= 1;
+//                cout << i << ',' << j << '=';
+//            }
+//            cout << endl;
+//        }
         dot(chunk_X, W, XW); // XW = X * W
-        cout << "after dot1\n";
+//        cout << "after dot1\n";
         add_to_each(XW, b, XWb); // XWb = X * W + b
 //        cout << "after add_to_each/**/\n";
         softmax(XWb, Y_prob); // Y_prob = softmax(X * W + b)
@@ -134,13 +142,14 @@ public:
                 cout << string_format("Epoch %d / %d", epoch, n_epochs) << endl;
 
             for (int batch_start = 0; batch_start < X.n_rows; batch_start += BATCH_SIZE) {
+//                cout << "batch start = " << X.data + batch_start * X.n_cols<< endl;
                 // TODO? make random batch generator
                 take_chunk(X, batch_start, chunk_X); // chunk_X = X[batch_start ... batch_start + CHUNK_SIZE]
 //                cout << "after take chunk X\n";
                 take_chunk(Y, batch_start, chunk_Y);
 //                cout << "after take chunk Y\n";
                 grad(); // grad_W and grad_b are now filled with result
-                cout << "after grad\n";
+//                cout << "after grad\n";
 
                 // TODO? regularization
                 scalar_mult(-lr, grad_W, lr_grad_W);
@@ -148,10 +157,10 @@ public:
 //                cout << "after scalar mults\n";
                 sub_from(W, lr_grad_W); // W = W - lr * grad_W
                 sub_from(b, lr_grad_b); // b = b - lr * grad_b
-                cout << "after update W & b\n";
+//                cout << "after update W & b\n";
 //                cost_history.push_back(cost); // TODO
             }
-            cout << "at end of epoch loop";
+//            cout << "at end of epoch loop";
 
             // TODO: add early stopping
             // Compute accuracy after each epoch

### 2017-02-08 evaluate.cpp
Update evaluate

@@ -19,21 +19,21 @@ double accuracy(const vector<int>& y_true, const vector<int>& y_pred) {
 // row-wise from_one_hot
 vector<int> from_one_hot_matrix(const Matrix& matrix) {
     vector<int> result;
-    result.reserve(n_rows(matrix));
+    result.reserve(matrix.n_rows);
 
-    for (const auto& row : matrix)
-        result.push_back(from_one_hot(row));
-
+    for (int i = 0; i < matrix.n_rows; ++i) {
+        result.push_back(from_one_hot(matrix.data[i], matrix.n_cols));
+    }
     return result;
 }
 
 // row-wise argmax
 vector<int> argmax_matrix(const Matrix& matrix) {
     vector<int> result;
-    result.reserve(n_rows(matrix));
+    result.reserve(matrix.n_rows);
 
-    for (const auto& row : matrix)
-        result.push_back(argmax(row));
-
+    for (int i = 0; i < matrix.n_rows; ++i) {
+        result.push_back(argmax(matrix.data[i], matrix.n_cols));
+    }
     return result;
 }

### 2017-02-08 main.cpp

@@ -58,11 +58,16 @@ int main(int argc, const char* argv[]) {
 //    double* w = v+2;
 //    cout << w[0];
 
-    int n_samples = 250;
+    int n_samples = 60000;
     Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/sample.csv"), images, labels);
+    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/sample.csv"), images, labels);
     NeuralNetwork net;
-    net.train(images, labels);
+    vector<double> cost_history, acc_history;
+    net.train(images, labels, cost_history,acc_history);
+    for (int i = 0; i < acc_history.size(); ++i) {
+        cout << acc_history[i] << endl;
+
+    }
 
 
 //    Matrix m(4, 2);

### 2017-02-08 nn.cpp

@@ -4,7 +4,8 @@
 #include <iostream>
 #include <fstream>
 #include "matrix.cpp"
-//#include "evaluate.cpp" // TODO
+#include "evaluate.cpp"
+#include <vector>
 
 
 using namespace std;
@@ -128,11 +129,11 @@ public:
     }
 
     void train(const Matrix& X, const Matrix& Y,
-//               vector<double>& cost_history, vector<double>& accuracy_history,
+               vector<double>& cost_history, vector<double>& accuracy_history,
                int n_epochs=10, int batch_size=100, double lr=0.1,
                bool verbose=true) {
-//        double cost, acc;
-//        auto Y_labels = from_one_hot_matrix(Y); // {5, 2, 9, ... }
+        double cost, acc;
+        vector<int> Y_labels = from_one_hot_matrix(Y); // {5, 2, 9, ... }
 
         random_init(W);
 //        cout << "after init W" << endl;
@@ -164,16 +165,18 @@ public:
 
             // TODO: add early stopping
             // Compute accuracy after each epoch
-//            acc = accuracy(predict(X), Y_labels); // TODO
-//            accuracy_history.push_back(acc);
+            acc = accuracy(predict(X), Y_labels);
+            accuracy_history.push_back(acc);
         }
 
     }
 
-//    vector<int> predict(const Matrix& X) {
-//        Matrix Y_prob = softmax(X * W + b);
-//        return argmax_matrix(Y_prob);
-//    }
+    vector<int> predict(const Matrix& X) {
+        dot(X, W, XW); // XW = X * W
+        add_to_each(XW, b, XWb); // XWb = X * W + b
+        softmax(XWb, Y_prob); // Y_prob = softmax(X * W + b)
+        return argmax_matrix(Y_prob);
+    }
 //
 //    int predict_one(const vector<double>& pixels, double& confidence) {
 //        // TODO: refactor to use .predict()

### 2017-02-08 utils.cpp

@@ -62,15 +62,15 @@ std::string string_format(const std::string fmt_str, ...) {
 
 
 // for 2 (and 5 classes) ~> {0 0 1 0 0}
-//vector<double> make_one_hot(int value, int n_classes=10) {
-//    vector<double> result(n_classes, 0);
-//    result[value] = 1;
-//    return result;
-//}
+vector<double> make_one_hot(int value, int n_classes=10) {
+    vector<double> result(n_classes, 0);
+    result[value] = 1;
+    return result;
+}
 
 // {0 0 1 0 0} ~> 2
-int from_one_hot(const vector<double>& v) {
-    for (int i = 0; i < v.size(); ++i)
+int from_one_hot(double* v, int n) {
+    for (int i = 0; i < n; ++i)
         if (v[i] == 1)
             return i;
 
@@ -78,11 +78,11 @@ int from_one_hot(const vector<double>& v) {
 }
 
 
-int argmax(vector<double> v) {
+int argmax(double* v, int n) {
     double max = v[0];
     int max_idx = 0;
 
-    for (int i = 1; i < v.size(); ++i)
+    for (int i = 1; i < n; ++i)
         if (v[i] > max) {
             max = v[i];
             max_idx = i;

### 2017-02-08 main.cpp
fix prediction

@@ -58,9 +58,9 @@ int main(int argc, const char* argv[]) {
 //    double* w = v+2;
 //    cout << w[0];
 
-    int n_samples = 60000;
+    int n_samples = 250;
     Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/sample.csv"), images, labels);
+    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/train.csv"), images, labels);
     NeuralNetwork net;
     vector<double> cost_history, acc_history;
     net.train(images, labels, cost_history,acc_history);
@@ -84,3 +84,4 @@ int main(int argc, const char* argv[]) {
 
     return 0;
 }
+

### 2017-02-08 nn.cpp

@@ -67,9 +67,9 @@ class NeuralNetwork {
     Matrix chunk_X = Matrix(CHUNK_SIZE, DATA_DIM);
     Matrix chunk_Y = Matrix(CHUNK_SIZE, N_CLASSES);
     // chunk sized
-        Matrix XW = Matrix(CHUNK_SIZE, N_CLASSES); // chunk_X * W ie (CHUNK_SIZE, DATA_DIM) * (DATA_DIM, N_CLASSES)
-        Matrix XWb = Matrix(CHUNK_SIZE, N_CLASSES);
-        Matrix Y_prob = Matrix(CHUNK_SIZE, N_CLASSES);
+        Matrix cXW = Matrix(CHUNK_SIZE, N_CLASSES); // chunk_X * W ie (CHUNK_SIZE, DATA_DIM) * (DATA_DIM, N_CLASSES)
+        Matrix cXWb = Matrix(CHUNK_SIZE, N_CLASSES);
+        Matrix cY_prob = Matrix(CHUNK_SIZE, N_CLASSES);
         Matrix delta = Matrix(CHUNK_SIZE, N_CLASSES);
 
         Matrix trX = Matrix(DATA_DIM, CHUNK_SIZE); // transpose CHUNK_SIZE, DATA_DIM
@@ -83,6 +83,11 @@ class NeuralNetwork {
     Matrix grad_W = Matrix(DATA_DIM, N_CLASSES); // like W
     Matrix grad_b = Matrix(1, N_CLASSES); // like b
 
+    // for prediction, on entire dataset
+    Matrix XW = Matrix(N_SAMPLES, N_CLASSES);
+    Matrix XWb = Matrix(N_SAMPLES, N_CLASSES);
+    Matrix Y_prob = Matrix(N_SAMPLES, N_CLASSES);
+
 
 public:
 
@@ -95,7 +100,7 @@ public:
     void grad() {
 //        print_dimensions(chunk_X);
 //        print_dimensions(W);
-//        print_dimensions(XW);
+//        print_dimensions(cXW);
 //        cout << "first elem of chunk x = " << chunk_X.data[0][0] << endl;
 //        for (int i = 0; i < chunk_X.n_rows; ++i) {
 //            for (int j = 0; j < chunk_X.n_cols; ++j) {
@@ -104,13 +109,13 @@ public:
 //            }
 //            cout << endl;
 //        }
-        dot(chunk_X, W, XW); // XW = X * W
[…]

### 2017-02-08 main.cpp
Edit add_to_each

@@ -60,7 +60,7 @@ int main(int argc, const char* argv[]) {
 
     int n_samples = 250;
     Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/train.csv"), images, labels);
+    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/sample.csv"), images, labels);
     NeuralNetwork net;
     vector<double> cost_history, acc_history;
     net.train(images, labels, cost_history,acc_history);

### 2017-02-08 matrix.cpp

@@ -66,7 +66,7 @@ void add_to_each(Matrix& matrix, const Matrix& to_add, Matrix& result) {
 
     for (int r = 0; r < to_add.n_rows; ++r)
         for (int c = 0; c < matrix.n_cols; ++c)
-            result.data[r][c] += to_add.data[0][c];
+            result.data[r][c] = matrix.data[r][c] + to_add.data[0][c];
 }
 
 void col_wise_sums(const Matrix& matrix, Matrix& result) {

### 2017-02-08 main.cpp
Fix add_to_each

@@ -58,7 +58,7 @@ int main(int argc, const char* argv[]) {
 //    double* w = v+2;
 //    cout << w[0];
 
-    int n_samples = 250;
+    int n_samples = 10000;
     Matrix images(n_samples, 784), labels(n_samples, 10);
     read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/sample.csv"), images, labels);
     NeuralNetwork net;

### 2017-02-08 matrix.cpp

@@ -56,7 +56,7 @@ std::istream& operator>>(std::istream& is, Matrix& matrix) {
 /*** Operators ***/
 // Adds first element of `to_add` to each element on the first column of `matrix`
 // second element of `to_add` to each of second column in `matrix`... etc
-void add_to_each(Matrix& matrix, const Matrix& to_add, Matrix& result) {
+void add_to_each(const Matrix& matrix, const Matrix& to_add, Matrix& result) {
     if (matrix.n_cols != to_add.n_cols)
         throw runtime_error(string_format("Add to each: matrix width and vector size are different: "
                                                   "%d, %d", matrix.n_rows, matrix.n_rows));
@@ -64,7 +64,7 @@ void add_to_each(Matrix& matrix, const Matrix& to_add, Matrix& result) {
         throw runtime_error(string_format("Add to each: vector to add has multiple rows: "
                                                   "%d", to_add.n_rows));
 
-    for (int r = 0; r < to_add.n_rows; ++r)
+    for (int r = 0; r < matrix.n_rows; ++r)
         for (int c = 0; c < matrix.n_cols; ++c)
             result.data[r][c] = matrix.data[r][c] + to_add.data[0][c];
 }

### 2017-02-08 nn.cpp

@@ -51,7 +51,7 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 
 class NeuralNetwork {
 
-    const int N_SAMPLES = 250;
+    const int N_SAMPLES = 10000;
     const int N_CLASSES = 10; // 10 digits from 0 to 9
     const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
     const int BATCH_SIZE = 100; // after how many pictures the weights are updated
@@ -183,7 +183,6 @@ public:
         dot(X, W, XW); // XW = X * W
         add_to_each(XW, b, XWb); // XWb = X * W + b
         softmax(XWb, Y_prob); // Y_prob = softmax(X * W + b)
-        cout << "here" << endl;
         return argmax_matrix(Y_prob);
     }
 //

### 2017-02-08 evaluate.cpp
Add stopwatch

@@ -16,24 +16,33 @@ double accuracy(const vector<int>& y_true, const vector<int>& y_pred) {
     return double(correct_preds) / n;
 }
 
-// row-wise from_one_hot
-vector<int> from_one_hot_matrix(const Matrix& matrix) {
-    vector<int> result;
-    result.reserve(matrix.n_rows);
+vector<int> labels_from_one_hot(const Matrix& matrix) {
+    vector<int> result(matrix.n_rows, 0);
 
-    for (int i = 0; i < matrix.n_rows; ++i) {
-        result.push_back(from_one_hot(matrix.data[i], matrix.n_cols));
-    }
+    for (int row = 0; row < matrix.n_rows; ++row)
+        for (int col = 0; col < matrix.n_cols; ++col) // {0 0 1 0 0} ~> 2
+            if (matrix.data[row][col] == 1) {
+                result[row] = col;
+                break;
+            }
+
     return result;
 }
 
-// row-wise argmax
-vector<int> argmax_matrix(const Matrix& matrix) {
-    vector<int> result;
-    result.reserve(matrix.n_rows);
+vector<int> argmax(const Matrix& matrix) {
+    vector<int> result(matrix.n_rows, 0);
+    double row_max;
+    int col_of_max;
 
-    for (int i = 0; i < matrix.n_rows; ++i) {
-        result.push_back(argmax(matrix.data[i], matrix.n_cols));
+    for (int row = 0; row < matrix.n_rows; ++row) {
+        double row_max = matrix.data[row][0];
+        for (int col = 1; col < matrix.n_cols; ++col) // {0 0 1 0 0} ~> 2
+            if (matrix.data[row][col] > row_max) {
+                row_max = matrix.data[row][col];
+                col_of_max = col;
+            }
+        result[row] = col_of_max;
     }
+
     return result;
 }

### 2017-02-08 main.cpp

@@ -2,6 +2,7 @@
 #include "reading.cpp"
 #include "nn.cpp"
 #include "matrix.cpp"
+#include <iomanip>
 
 using namespace std;
 
@@ -60,14 +61,14 @@ int main(int argc, const char* argv[]) {
 
     int n_samples = 10000;
     Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/sample.csv"), images, labels);
+    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/test.csv"), images, labels);
     NeuralNetwork net;
     vector<double> cost_history, acc_history;
-    net.train(images, labels, cost_history,acc_history);
-    for (int i = 0; i < acc_history.size(); ++i) {
-        cout << acc_history[i] << endl;
-
-    }
+    cout << setprecision(3);
+    net.train(images, labels, cost_history, acc_history);
+//    for (int i = 0; i < acc_history.size(); ++i) {
+//        cout << acc_history[i] << endl;
+//    }
 
 
 //    Matrix m(4, 2);

### 2017-02-08 nn.cpp

@@ -6,6 +6,7 @@
 #include "matrix.cpp"
 #include "evaluate.cpp"
 #include <vector>
+#include <chrono>
 
 
 using namespace std;
@@ -52,10 +53,11 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 class NeuralNetwork {
 
     const int N_SAMPLES = 10000;
+    const int N_WORKERS = 1; // how many processes compute gradient in parallel
+
     const int N_CLASSES = 10; // 10 digits from 0 to 9
     const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
     const int BATCH_SIZE = 100; // after how many pictures the weights are updated
-    const int N_WORKERS = 4; // how many processes compute gradient in parallel
     const int CHUNK_SIZE = BATCH_SIZE / N_WORKERS; // how many rows each worker gets
 
     // X dimensions: N_SAMPLES x DATA_DIM
@@ -98,25 +100,10 @@ public:
 //    }
 
     void grad() {
-//        print_dimensions(chunk_X);
-//        print_dimensions(W);
-//        print_dimensions(cXW);
-//        cout << "first elem of chunk x = " << chunk_X.data[0][0] << endl;
-//        for (int i = 0; i < chunk_X.n_rows; ++i) {
-//            for (int j = 0; j < chunk_X.n_cols; ++j) {
-//                chunk_X.data[i][j] *= 1;
-//                cout << i << ',' << j << '=';
-//            }
-//            cout << endl;
-//        }
         dot(chunk_X, W, cXW); // cXW = X * W
-//        cout << "after dot1\n";
         add_to_each(cXW, b, cXWb); // cXWb = X * W + b
-//        cout << "after add_to_each/**/\n";
         softmax(cXWb, cY_prob); // cY_prob = softmax(X * W + b)
-//        cout << "after softmax\n";
         sub(cY_prob, chunk_Y, delta); // delta = cY_prob - Y
-//        cout << "after sub1\n";
 
         double contribution = 1/double(CHUNK_SIZE);
 
@@ -135,55 +122,51 @@ public:
 
     void train(const Matrix& X, const Matrix& Y,
                vector<double>& cost_history, vector<double>& accuracy_history,
-               int n_epochs=10, int batch_size=100, double lr=0.1,
+               int n_epochs=100, double lr=0.5,
                bool verbose=true) {
         double cost, acc;
[…]

### 2017-02-08 utils.cpp

@@ -68,25 +68,3 @@ vector<double> make_one_hot(int value, int n_classes=10) {
     return result;
 }
 
-// {0 0 1 0 0} ~> 2
-int from_one_hot(double* v, int n) {
-    for (int i = 0; i < n; ++i)
-        if (v[i] == 1)
-            return i;
-
-    throw runtime_error("Vector to convert from one-hot form contains no ones");
-}
-
-
-int argmax(double* v, int n) {
-    double max = v[0];
-    int max_idx = 0;
-
-    for (int i = 1; i < n; ++i)
-        if (v[i] > max) {
-            max = v[i];
-            max_idx = i;
-        }
-
-    return max_idx;
-}

### 2017-02-08 nn.cpp
amend

@@ -156,8 +156,8 @@ public:
             // TODO: add early stopping
             // Compute accuracy after each epoch
             acc = accuracy(predict(X), Y_labels);
-            cout << "accuracy: " << acc * 100 << "%";
-            accuracy_history.push_back(acc) << endl;
+            cout << "accuracy: " << acc * 100 << "%" << endl;
+            accuracy_history.push_back(acc);
         }
 
     }

### 2017-02-08 main.cpp
start to work on mpi

@@ -59,9 +59,9 @@ int main(int argc, const char* argv[]) {
 //    double* w = v+2;
 //    cout << w[0];
 
-    int n_samples = 10000;
+    int n_samples = 250;
     Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/test.csv"), images, labels);
+    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/sample.csv"), images, labels, false);
     NeuralNetwork net;
     vector<double> cost_history, acc_history;
     cout << setprecision(3);

### 2017-02-08 matrix.cpp

@@ -8,11 +8,13 @@ class Matrix {
 
 public:
     double** data;
-    int n_rows;
-    int n_cols;
+    const int n_rows;
+    const int n_cols;
+    const int n_elements;
     bool is_chunk = false;
 
-    Matrix(int n_rows,int n_cols): n_rows(n_rows), n_cols(n_cols) {
+    Matrix(int n_rows,int n_cols):
+            n_rows(n_rows), n_cols(n_cols), n_elements(n_rows * n_cols) {
         data = new double*[n_rows];
         for (int i = 0; i < n_rows; ++i)
             data[i] = new double[n_cols];
@@ -31,6 +33,7 @@ public:
             for (int j = 0; j < n_cols; ++j)
                 data[i][j] = 0;
     }
+
 };
 
 void print_dimensions(const Matrix& m) {
@@ -138,16 +141,16 @@ void sub(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
             result.data[i][j] = rhs.data[i][j] - lhs.data[i][j];
 }
 
-void sub_from(Matrix& from, const Matrix& to_sub) {
-    // from = from - to_sub
-    if ((from.n_rows != to_sub.n_rows) || (from.n_cols != to_sub.n_cols))
+void add_to(Matrix& to, const Matrix& from) {
+    // to = to + from
+    if ((to.n_rows != from.n_rows) || (to.n_cols != from.n_cols))
         throw runtime_error(string_format("Matrix FROM subtraction: number of rows/cols is different: "
-                                                  "from = (%d, %d), to_sub = (%d, %d)",
-                                          from.n_rows, from.n_cols, to_sub.n_rows, to_sub.n_cols));
+                                                  "to = (%d, %d), from = (%d, %d)",
+                                          to.n_rows, to.n_cols, from.n_rows, from.n_cols));
 
-    for (int i = 0; i < from.n_rows; ++i)
-        for (int j = 0; j < from.n_cols; ++j)
-            from.data[i][j] -= to_sub.data[i][j];
+    for (int i = 0; i < to.n_rows; ++i)
+        for (int j = 0; j < to.n_cols; ++j)
+            to.data[i][j] += from.data[i][j];
 }
 
 

### 2017-02-08 nn.cpp

@@ -7,9 +7,17 @@
 #include "evaluate.cpp"
 #include <vector>
 #include <chrono>
+#include "mpi.h"
 
 
 using namespace std;
+
+
+/*** MPI constants ***/
+const int MASTER = 0;
+const int GRAD_W_TAG    = 0;
+const int GRAD_B_TAG    = 1;
+
 
 
 /*** Requisites ***/
@@ -52,12 +60,12 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 
 class NeuralNetwork {
 
-    const int N_SAMPLES = 10000;
-    const int N_WORKERS = 1; // how many processes compute gradient in parallel
+    const int N_SAMPLES = 250;
+    const int N_WORKERS = 2; // how many processes compute gradient in parallel
 
     const int N_CLASSES = 10; // 10 digits from 0 to 9
     const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
-    const int BATCH_SIZE = 100; // after how many pictures the weights are updated
+    const int BATCH_SIZE = 50; // after how many pictures the weights are updated
     const int CHUNK_SIZE = BATCH_SIZE / N_WORKERS; // how many rows each worker gets
 
     // X dimensions: N_SAMPLES x DATA_DIM
@@ -82,8 +90,12 @@ class NeuralNetwork {
 
     Matrix lr_grad_W = Matrix(DATA_DIM, N_CLASSES); // like W
     Matrix lr_grad_b = Matrix(1, N_CLASSES); // like b
-    Matrix grad_W = Matrix(DATA_DIM, N_CLASSES); // like W
-    Matrix grad_b = Matrix(1, N_CLASSES); // like b
+
+    Matrix partial_grad_W = Matrix(DATA_DIM, N_CLASSES); // like W
+    Matrix partial_grad_b = Matrix(1, N_CLASSES); // like b
+
+    Matrix total_grad_W = Matrix(DATA_DIM, N_CLASSES); // like W
+    Matrix total_grad_b = Matrix(1, N_CLASSES); // like b
 
     // for prediction, on entire dataset
     Matrix XW = Matrix(N_SAMPLES, N_CLASSES);
@@ -108,13 +120,11 @@ public:
         double contribution = 1/double(CHUNK_SIZE);
 
         transpose(chunk_X, trX); // trX = transpose(X)
-//        cout << "after transpose\n";
         dot(trX, delta, trXdelta); // trXdelta = transpose(X) * delta
[…]

### 2017-02-08 matrix.cpp
Fix substract error

@@ -139,7 +139,7 @@ void sub(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
 
     for (int i = 0; i < rhs.n_rows; ++i)
         for (int j = 0; j < rhs.n_cols; ++j)
-            result.data[i][j] = rhs.data[i][j] - lhs.data[i][j];
+            result.data[i][j] = lhs.data[i][j] - rhs.data[i][j];
 }
 
 void add_to(Matrix& to, const Matrix& from) {

### 2017-02-08 main.cpp
Add parralel transfer without parameter bcast

@@ -61,11 +61,11 @@ int main(int argc, const char* argv[]) {
 
     int n_samples = 250;
     Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/sample.csv"), images, labels, false);
+    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/sample.csv"), images, labels, false);
     NeuralNetwork net;
     vector<double> cost_history, acc_history;
     cout << setprecision(3);
-    net.train(images, labels, cost_history, acc_history);
+    net.train(images, labels, cost_history, acc_history, 5);
 //    for (int i = 0; i < acc_history.size(); ++i) {
 //        cout << acc_history[i] << endl;
 //    }

### 2017-02-08 matrix.cpp

@@ -141,11 +141,22 @@ void sub(const Matrix& lhs, const Matrix& rhs, Matrix& result) {
         for (int j = 0; j < rhs.n_cols; ++j)
             result.data[i][j] = lhs.data[i][j] - rhs.data[i][j];
 }
+void sub_from(Matrix& to, const Matrix& from) {
+    // to = to + from
+    if ((to.n_rows != from.n_rows) || (to.n_cols != from.n_cols))
+        throw runtime_error(string_format("Matrix FROM subtraction: number of rows/cols is different: "
+                                                  "to = (%d, %d), from = (%d, %d)",
+                                          to.n_rows, to.n_cols, from.n_rows, from.n_cols));
+
+    for (int i = 0; i < to.n_rows; ++i)
+        for (int j = 0; j < to.n_cols; ++j)
+            to.data[i][j] -= from.data[i][j];
+}
 
 void add_to(Matrix& to, const Matrix& from) {
     // to = to + from
     if ((to.n_rows != from.n_rows) || (to.n_cols != from.n_cols))
-        throw runtime_error(string_format("Matrix FROM subtraction: number of rows/cols is different: "
+        throw runtime_error(string_format("Matrix FROM addition: number of rows/cols is different: "
                                                   "to = (%d, %d), from = (%d, %d)",
                                           to.n_rows, to.n_cols, from.n_rows, from.n_cols));
 

### 2017-02-08 nn.cpp

@@ -61,7 +61,7 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 class NeuralNetwork {
 
     const int N_SAMPLES = 250;
-    const int N_WORKERS = 2; // how many processes compute gradient in parallel
+    const int N_WORKERS = 1; // how many processes compute gradient in parallel
 
     const int N_CLASSES = 10; // 10 digits from 0 to 9
     const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
@@ -165,48 +165,49 @@ public:
 
                     // Get partial gradients from each worker
                     for (int worker = 1; worker <= N_WORKERS; ++worker) {
-//                        cout << "master: waiting for worker #" << worker << " in epoch " << epoch << endl;
+//                        cout << partial_grad_W.data[0] << " " << &(partial_grad_W.data[0][0]) << endl;
+                        cout << "master: waiting for worker #" << worker << " in epoch " << epoch << endl;
                         cout << ">>>>>master: epoch " << epoch << ", batch_start " << batch_start << ", worker #" << worker <<  endl;
-                        MPI_Recv(&(partial_grad_W.data[0][0]), partial_grad_W.n_elements, MPI_DOUBLE, worker, GRAD_W_TAG, MPI_COMM_WORLD, MPI_STATUS_IGNORE);
-                        MPI_Recv(&(partial_grad_b.data[0][0]), partial_grad_b.n_elements, MPI_DOUBLE, worker, GRAD_B_TAG, MPI_COMM_WORLD, MPI_STATUS_IGNORE);
-//                        cout << "here" << endl;
-//                        cout << partial_grad_b.data[0][0] << endl;
+                        MPI_Recv(partial_grad_W.data[0], partial_grad_W.n_elements, MPI_DOUBLE,
+                                 worker, GRAD_W_TAG, MPI_COMM_WORLD, MPI_STATUS_IGNORE);
+                        MPI_Recv(partial_grad_b.data[0], partial_grad_b.n_elements, MPI_DOUBLE,
+                                 worker, GRAD_B_TAG, MPI_COMM_WORLD, MPI_STATUS_IGNORE);
+                        cout << "master: After Recv for worker #" << worker << " in epoch " << epoch << endl;
+                        // Sum partial gradients
+                        add_to(total_grad_W, partial_grad_W);
+                        add_to(total_grad_b, partial_grad_b);
                     }
[…]

### 2017-02-08 main.cpp
Update: bcast params

@@ -59,13 +59,13 @@ int main(int argc, const char* argv[]) {
 //    double* w = v+2;
 //    cout << w[0];
 
-    int n_samples = 250;
+    int n_samples = 10000;
     Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/sample.csv"), images, labels, false);
+    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/test.csv"), images, labels, false);
     NeuralNetwork net;
     vector<double> cost_history, acc_history;
     cout << setprecision(3);
-    net.train(images, labels, cost_history, acc_history, 5);
+    net.train(images, labels, cost_history, acc_history, 10);
 //    for (int i = 0; i < acc_history.size(); ++i) {
 //        cout << acc_history[i] << endl;
 //    }

### 2017-02-08 nn.cpp

@@ -60,12 +60,12 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 
 class NeuralNetwork {
 
-    const int N_SAMPLES = 250;
+    const int N_SAMPLES = 10000;
     const int N_WORKERS = 1; // how many processes compute gradient in parallel
 
     const int N_CLASSES = 10; // 10 digits from 0 to 9
     const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
-    const int BATCH_SIZE = 50; // after how many pictures the weights are updated
+    const int BATCH_SIZE = 1000; // after how many pictures the weights are updated
     const int CHUNK_SIZE = BATCH_SIZE / N_WORKERS; // how many rows each worker gets
 
     // X dimensions: N_SAMPLES x DATA_DIM
@@ -161,45 +161,36 @@ public:
                     // clear summed from previous batch
                     total_grad_W.clear();
                     total_grad_b.clear();
-//                    cout << ">>>>>master: epoch " << epoch << ", batch_start " << batch_start <<  endl;
 
                     // Get partial gradients from each worker
                     for (int worker = 1; worker <= N_WORKERS; ++worker) {
-//                        cout << partial_grad_W.data[0] << " " << &(partial_grad_W.data[0][0]) << endl;
                         cout << "master: waiting for worker #" << worker << " in epoch " << epoch << endl;
-                        cout << ">>>>>master: epoch " << epoch << ", batch_start " << batch_start << ", worker #" << worker <<  endl;
                         MPI_Recv(partial_grad_W.data[0], partial_grad_W.n_elements, MPI_DOUBLE,
                                  worker, GRAD_W_TAG, MPI_COMM_WORLD, MPI_STATUS_IGNORE);
                         MPI_Recv(partial_grad_b.data[0], partial_grad_b.n_elements, MPI_DOUBLE,
                                  worker, GRAD_B_TAG, MPI_COMM_WORLD, MPI_STATUS_IGNORE);
-                        cout << "master: After Recv for worker #" << worker << " in epoch " << epoch << endl;
                         // Sum partial gradients
                         add_to(total_grad_W, partial_grad_W);
                         add_to(total_grad_b, partial_grad_b);
                     }
 
-                    // After receivtake_chunk(X, start_index, chunk_X); // chunk_X = X[batch_start ... batch_start + CHUNK_SIZE]
+                    // After receive
[…]

### 2017-02-08 main.cpp
Update nn, compute accuracy on master

@@ -59,13 +59,16 @@ int main(int argc, const char* argv[]) {
 //    double* w = v+2;
 //    cout << w[0];
 
-    int n_samples = 10000;
+    int n_samples = 60000;
     Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/test.csv"), images, labels, false);
+    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/train.csv"), images, labels, false);
     NeuralNetwork net;
     vector<double> cost_history, acc_history;
     cout << setprecision(3);
-    net.train(images, labels, cost_history, acc_history, 10);
+    net.train(images, labels, cost_history, acc_history, 10, 0.5);
+    for(auto acc : acc_history) {
+        cout << acc << endl;
+    }
 //    for (int i = 0; i < acc_history.size(); ++i) {
 //        cout << acc_history[i] << endl;
 //    }

### 2017-02-08 nn.cpp

@@ -60,12 +60,12 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 
 class NeuralNetwork {
 
-    const int N_SAMPLES = 10000;
-    const int N_WORKERS = 1; // how many processes compute gradient in parallel
+    const int N_SAMPLES = 60000;
+    const int N_WORKERS = 6; // how many processes compute gradient in parallel
 
     const int N_CLASSES = 10; // 10 digits from 0 to 9
     const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
-    const int BATCH_SIZE = 1000; // after how many pictures the weights are updated
+    const int BATCH_SIZE = 600; // after how many pictures the weights are updated
     const int CHUNK_SIZE = BATCH_SIZE / N_WORKERS; // how many rows each worker gets
 
     // X dimensions: N_SAMPLES x DATA_DIM
@@ -132,7 +132,7 @@ public:
 
     void train(const Matrix& X, const Matrix& Y,
                vector<double>& cost_history, vector<double>& accuracy_history,
-               int n_epochs=100, double lr=0.5,
+               int n_epochs=100, double lr=0.1,
                bool verbose=true) {
         if (N_SAMPLES != X.n_rows) throw runtime_error("bad N_SAMPLES");
 
@@ -151,7 +151,7 @@ public:
 
         for (int epoch = 1; epoch <= n_epochs; ++epoch) {
             if (rank == MASTER) {
-//                cout << string_format("Epoch %d / %d", epoch, n_epochs) << flush;
+                cout << string_format("Epoch %d / %d", epoch, n_epochs) << flush;
                 start_time = std::chrono::system_clock::now();
             }
 
@@ -164,7 +164,7 @@ public:
 
                     // Get partial gradients from each worker
                     for (int worker = 1; worker <= N_WORKERS; ++worker) {
-                        cout << "master: waiting for worker #" << worker << " in epoch " << epoch << endl;
+//                        cout << "master: waiting for worker #" << worker << " in epoch " << epoch << endl;
                         MPI_Recv(partial_grad_W.data[0], partial_grad_W.n_elements, MPI_DOUBLE,
                                  worker, GRAD_W_TAG, MPI_COMM_WORLD, MPI_STATUS_IGNORE);
                         MPI_Recv(partial_grad_b.data[0], partial_grad_b.n_elements, MPI_DOUBLE,
@@ -188,7 +188,7 @@ public:
                 if (rank != MASTER) { // worker
                     // TODO? make random batch generator
[…]

### 2017-02-08 doc/Paper.tex
Add plots, b&w model graph

@@
 
+\begin{center}
 \begin{tikzpicture}
-\centering
 \begin{axis} [
@@
   legend pos=south east,
+  axis line style={black!65, line width=.1pt}, % thinner frame border
+  legend style={draw=black!50, line width=.1pt}, % thinner frame border
+  every x tick/.style={white}, % hide x ticks (make them white)
+  every y tick/.style={white}, % hide y ticks (make them white)
+  every axis plot/.append style={thick},
 ]
@@
   };
-\addplot[dashed]
+\addplot[dash pattern=on 6pt off 4pt]
   coordinates {
@@
 
-\legend{train set, test set}
+\legend{train, test}
 \end{axis}
 \end{tikzpicture}
+\end{center}
 
@@
 TODO: graph of train time vs nr of processors
+
+\begin{center}
+\begin{tikzpicture}
+\begin{axis}[
+  title={Time performance},
+  xlabel=Processes,
+  ylabel=Seconds,
+  bar width=.8cm,
+  axis line style={black!65, line width=.1pt}, % thinner border
+  symbolic x coords={1,2,4,6,8,12,16}, % equidistant bars
+  every x tick/.style={white}, % hide x ticks (make them white)
+  every y tick/.style={line width=.1pt, draw=gray!10}, % y ticks same color as grid
+  ymajorgrids=true, % show horizontal grid lines
+  grid style={line width=.1pt, draw=gray!10}, % gray grid lines
+  ]
+  \addplot[ybar, black!40, fill=black!20] coordinates {
+      (1,8) 
+      (2,5)
+      (4,4) 
+      (6,3) 
+      (8,2.5) 
+      (12,2) 
+      (16,1.5)    
+      };
+\end{axis}
+\end{tikzpicture}
+\end{center}
 

### 2017-02-08 doc/Paper.tex
Continue parallelization section

@@
 \section{Parallelization}
-Parallelism is needed in the training phase of the model, as prediction happens instantaneously
+The phase that needs parallelization is the training of the model.
+Predicting is not computationally intensive.
+The bulk of the time is spent iterating over the entire training set multiple times.
+In one epoch, the network looks at every image, in batches and does the following:
+\begin{itemize}
+\item compute the change that must happen in order to better recognize -- called gradient,
+\item updates its internal weights based on the gradient.
+\end{itemize}
+Computing the gradient takes the most time. The expensive operations are two dot-products on the weights matrices and the a matrix-wise softmax. This is the portion that we will parallelize.
+
+When looking at a batch of images:
+\begin{itemize}
+\item the batch is assigned evenly among workers,
+\item each worker computes the gradient on its chunk of the batch,
+\item partial gradients are sent from each worker to the master,
+\item the master process accumulates all the partial gradients and aggregates them,
+\item the master updates the model weights based on the total gradient,
+\item updated weights are sent back to every worker,
+\item a new batch is ready to be processed: each worker now knows the effects of all the others.
+\end{itemize}
+
+In our case, aggregating the partial gradients requires nothing more then a summation.
+\\
+
+Naturally, there appears a running time -- model accuracy trade-off.
+Updating the weights after every image yields better accuracy than updating once every 1,000 samples.
+In the case of a greater batch size, the model has not evolved after looking at the first image, nor after the second or the third.
+It still analyzes them knowing just as much as it did when looking at the first sample.
+
+In contrast, a greater batch size produces better parallelization performance.
+If each worker is assigned a single image, the communication cost of transferring gradient and weight matrices far outweighs the time taken to compute them.
+Efficiency improvement is present only when the time taken to transfer the results is lower than the time taken to compute them.
+Greater batch sizes incur more infrequent updates thus reducing the communication overhead.
+
+The trade-off is now clear: a lower batch size leads to greater accuracy while a greater batch size causes faster training.
 
 TODO: talk about matrix multiplication overhead
-
-TODO: graph of train time vs nr of processors
 
@@
 \begin{axis}[
-  title={Time performance},
+  title={Time per epoch},
   xlabel=Processes,

### 2017-02-08 doc/Paper.tex
Explain 2D-array choice in Implementation section

@@
 The trade-off is now clear: a lower batch size leads to greater accuracy while a greater batch size causes faster training.
+\\
+
+Another
 
@@
 This includes the neural network training and prediction.
-Data is maintained in memory using vectors from the standard library.
+Model weights and image samples are kept in memory using 2D-arrays.
+
+Matrices build as vector of vectors (using the standard library implementation), while providing greater flexibility, cannot be sent efficiently using MPI.
+MPI deals with contiguous blocks of memory, while STL vectors are, by definition, dynamic.
+A workaround this is to send a matrix row-by-row.
+Our initial implementation used this method but unfortunately it provides little to no increase in parallelization performance and caused a general-rewrite.
 

### 2017-02-08 doc/Paper.tex
Talk about matrix multiplication

@@
 	\maketitle
-	\setlength{\parindent}{0pt}
+% 	\setlength{\parindent}{0pt}
 
@@
 
+\begin{figure}[htb]
+\centering
+\includegraphics[width=8cm]{images/model-graph.png}
+\caption{Simplified representation of model graph \cite{tf}}
+\end{figure}
+
 The cost function is cross-entropy:
@@
 
-\begin{figure}[htb]
-\centering
-\includegraphics[width=8cm]{images/model-graph.png}
-\caption{Simplified representation of model graph \cite{tf}}
-\end{figure}
-
-\paragraph{Training} We employ gradient-descent on mini-batches. Early-stopping prevents the model from further learning when performance on the validation split starts decreasing.
+\paragraph{Training} We employ gradient-descent on mini-batches.
+Early-stopping prevents the model from further learning when performance on the validation split starts decreasing.
+Random sampling averts cycles when iterating through the dataset.
 
@@
 \paragraph{Evaluation}
+Model performance is measured on a separate testing set using the accuracy metric -- ie: the percentage of images correctly classified.
+
 TODO
@@
 \end{itemize}
-
 In our case, aggregating the partial gradients requires nothing more then a summation.
@@
 
-Another
-
-TODO: talk about matrix multiplication overhead
+Another approach would be to let the algorithm run sequentially and parallelize only the matrix multiplication operation.
+In light of the previously presented considerations, we can conclude that this would be grossly inefficient.
+The transfer overhead is much higher than the benefit gained by splitting the problem.
 
@@
 MPI deals with contiguous blocks of memory, while STL vectors are, by definition, dynamic.
-A workaround this is to send a matrix row-by-row.
+A work-around for this is to send a matrix row-by-row.
 Our initial implementation used this method but unfortunately it provides little to no increase in parallelization performance and caused a general-rewrite.
+
+The most expensive matrix operations are dot products and exponentiation.
+For matrix multiplication, we explored the Strassen algorithm, $O(n^{2.8})$, and the Coppersmith–Winograd algorithm, $O(n^{2.37})$.
+Although they achieve high asymptotic performance, hidden constants make them impractical in our case.
+We went with the Naive, $\Theta(n^3)$, algorithm with the $ikj$ optimization.
 
@@
 
-The entire source code is available at \url{https://github.com/stefan1niculae/mnist-in-mpi}.
+\noindent The entire source code is available at \url{https://github.com/stefan1niculae/mnist-in-mpi}.
 

### 2017-02-08 main.cpp
Add cost function

@@ -59,13 +59,13 @@ int main(int argc, const char* argv[]) {
 //    double* w = v+2;
 //    cout << w[0];
 
-    int n_samples = 60000;
+    int n_samples = 10000;
     Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/train.csv"), images, labels, false);
+    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/test.csv"), images, labels, false);
     NeuralNetwork net;
     vector<double> cost_history, acc_history;
     cout << setprecision(3);
-    net.train(images, labels, cost_history, acc_history, 10, 0.5);
+    net.train(images, labels, cost_history, acc_history, 100, 0.5);
     for(auto acc : acc_history) {
         cout << acc << endl;
     }

### 2017-02-08 matrix.cpp

@@ -107,13 +107,14 @@ void log(const Matrix& matrix, Matrix& result) {
 }
 
 // whole matrix sum
-void sum(const Matrix& matrix, double result) {
-    result = 0;
+double sum(const Matrix& matrix) {
+    double result = 0;
     for (int i = 0; i < matrix.n_rows ; ++i) {
         for (int j = 0; j < matrix.n_cols; ++j) {
             result += matrix.data[i][j];
         }
     }
+    return result;
 }
 
 //

### 2017-02-08 nn.cpp

@@ -17,6 +17,7 @@ using namespace std;
 const int MASTER = 0;
 const int GRAD_W_TAG    = 0;
 const int GRAD_B_TAG    = 1;
+const int COST_TAG      = 2;
 
 
 
@@ -39,12 +40,14 @@ void softmax(const Matrix& matrix, Matrix& result) {
     }
 }
 
-// TODO
-//double cross_entropy(const Matrix& Y, const Matrix& Y_prob) {
-//    // Cost function
-//    Matrix result = hadamard(Y, log(Y_prob));
-//    return 1/double(n_rows(Y)) * sum(result);
-//}
+double cross_entropy(const Matrix& Y, const Matrix& Y_prob) {
+    // Cost function
+    Matrix logY_prob(Y.n_rows, Y.n_cols);
+    Matrix result(Y.n_rows, Y.n_cols);
+    log(Y_prob, logY_prob);
+    hadamard(Y, logY_prob, result);
+    return -1/double(Y.n_rows) * sum(result);
+}
 
 void random_init(Matrix& W, double mean=0., double std=1.) {
     default_random_engine generator;
@@ -60,12 +63,12 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 
 class NeuralNetwork {
 
-    const int N_SAMPLES = 60000;
-    const int N_WORKERS = 6; // how many processes compute gradient in parallel
+    const int N_SAMPLES = 10000;
+    const int N_WORKERS = 5; // how many processes compute gradient in parallel
 
     const int N_CLASSES = 10; // 10 digits from 0 to 9
     const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
-    const int BATCH_SIZE = 600; // after how many pictures the weights are updated
+    const int BATCH_SIZE = 500; // after how many pictures the weights are updated
     const int CHUNK_SIZE = BATCH_SIZE / N_WORKERS; // how many rows each worker gets
 
     // X dimensions: N_SAMPLES x DATA_DIM
@@ -111,7 +114,7 @@ public:
 //        this->load(path);
 //    }
 
-    void grad() {
+    double grad() {
         dot(chunk_X, W, cXW); // cXW = X * W
         add_to_each(cXW, b, cXWb); // cXWb = X * W + b
         softmax(cXWb, cY_prob); // cY_prob = softmax(X * W + b)
@@ -127,7 +130,7 @@ public:
[…]

### 2017-02-08 doc/Paper.tex
Paper: more visible TODOs & interactive section

@@
   \tikzexternalize
+\usepackage{todonotes}
 
@@
   \setheadsepline[122mm]{0.3mm}
+
+% floor operator
+\newcommand{\floor}[1]{\lfloor {#1} \rfloor}
+\newcommand{\TODO}[1]{\noindent \textbf{\textcolor{red}{TODO: #1}}}
 
@@
 \begin{abstract}
-TODO: rewrite when paper is done, talk about each section
+\TODO{rewrite when paper is done, talk about each section}
 
@@
 \newpage
-TODO: introduction?
+\TODO{introduction?}
 
@@
 Given a picture of a handwritten digit, label it accordingly 0 - 9.
+\TODO{more here}
 
@@
 \paragraph{Prediction}
-TODO
+\TODO{explain prediction?}
 
@@
 
-TODO
+\TODO{more about evaluation?}
+
+\noindent The following chart illustrates model evolution, both on the test and on the training set, after each pass through the dataset.
 
@@
 \end{center}
+
+\TODO{comment on chart}
 
@@
 
+The following chart illustrates the improvement brought by parallelization on one pass through the data:
+
 \begin{center}
@@
 
+\TODO{comment on chart}
+
 \newpage
 \section{Graphical Interface}
-TODO: picture of resizing
+The user is presented with a blank box in which to draw a digit to be recognized. Upon request, the prediction along with the model's confidence in it is displayed.
 
-TODO: picture of GUI
+\missingfigure[figcolor=white]{GUI}
+
+\noindent Since the model was trained on $28 \times 28$ images, we need to pass the input in the same format.
+On today's screens, an input box of $28 \times 28$ pixels would look comically small.
+The raw input must be resized accordingly.
+
+The simplest down-sizing method is taking each block of pixels in the original image and averaging it to form a single pixel in the resulting image.
+For simplicity, we assume square images and pixels with only one channel -- grayscale.
+Eg: an image with $n^2$ original pixels is condensed to $s^2$ pixels, meaning a factor of $f=\frac{n}{s}$.
+That means each of the $s^2$ pixels is a sum of a $f \times f$ block divided by $f^2$.
+
+The trouble appears with encoding the image -- both the raw image and the model's inputs are in array form instead of a matrix.
+Ie: the pixels are all placed one after another starting from the top left corner.
+
+\missingfigure[figcolor=white]{resizing example pdf}
+The transformation can be computed by \textit{compacting} the columns:
+$\sum_i Orig_i$ for column at index $\floor{\frac{i}{f}}$.
+Then for rows: $\sum_i Col_i$ for pixel at index $(i \bmod s) + \floor{\frac{i}{s \times f}}$.
+Where $Orig$ denotes the the original image $Col$ the column-wise averages.
 
@@
 We went with the Naive, $\Theta(n^3)$, algorithm with the $ikj$ optimization.
+For exponentiation, we rely on STL's implementation, although this can be further optimized.
 

### 2017-02-08 main.cpp
Dynamic matrix allocation

@@ -59,13 +59,13 @@ int main(int argc, const char* argv[]) {
 //    double* w = v+2;
 //    cout << w[0];
 
-    int n_samples = 10000;
+    int n_samples = 2000;
     Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/home/ionut/workspace/ppc/mnist-in-mpi/data/test.csv"), images, labels, false);
+    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/test.csv"), images, labels, false);
     NeuralNetwork net;
     vector<double> cost_history, acc_history;
     cout << setprecision(3);
-    net.train(images, labels, cost_history, acc_history, 100, 0.5);
+    net.train(images, labels, cost_history, acc_history, 100, 200, 0.5);
     for(auto acc : acc_history) {
         cout << acc << endl;
     }

### 2017-02-08 matrix.cpp

@@ -13,7 +13,7 @@ public:
     const int n_elements;
     bool is_chunk = false;
 
-    Matrix(int n_rows,int n_cols):
+    Matrix(int n_rows,int n_cols) :
             n_rows(n_rows), n_cols(n_cols), n_elements(n_rows * n_cols) {
 
         double* mem = (double*)malloc(n_rows * n_cols * sizeof(double));

### 2017-02-08 nn.cpp

@@ -19,6 +19,8 @@ const int GRAD_W_TAG    = 0;
 const int GRAD_B_TAG    = 1;
 const int COST_TAG      = 2;
 
+static vector<double> NONE = vector<double>();
+
 
 
 /*** Requisites ***/
@@ -63,13 +65,9 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 
 class NeuralNetwork {
 
-    const int N_SAMPLES = 10000;
-    const int N_WORKERS = 5; // how many processes compute gradient in parallel
-
     const int N_CLASSES = 10; // 10 digits from 0 to 9
     const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
-    const int BATCH_SIZE = 500; // after how many pictures the weights are updated
-    const int CHUNK_SIZE = BATCH_SIZE / N_WORKERS; // how many rows each worker gets
+
 
     // X dimensions: N_SAMPLES x DATA_DIM
     // Y dimensions: N_SAMPLES x N_CLASSES
@@ -77,33 +75,6 @@ class NeuralNetwork {
     Matrix W = Matrix(DATA_DIM, N_CLASSES);        // TODO? more layers
     Matrix b = Matrix(1, N_CLASSES);
 
-    Matrix chunk_X = Matrix(CHUNK_SIZE, DATA_DIM);
-    Matrix chunk_Y = Matrix(CHUNK_SIZE, N_CLASSES);
-    // chunk sized
-        Matrix cXW = Matrix(CHUNK_SIZE, N_CLASSES); // chunk_X * W ie (CHUNK_SIZE, DATA_DIM) * (DATA_DIM, N_CLASSES)
-        Matrix cXWb = Matrix(CHUNK_SIZE, N_CLASSES);
-        Matrix cY_prob = Matrix(CHUNK_SIZE, N_CLASSES);
-        Matrix delta = Matrix(CHUNK_SIZE, N_CLASSES);
-
-        Matrix trX = Matrix(DATA_DIM, CHUNK_SIZE); // transpose CHUNK_SIZE, DATA_DIM
-        Matrix trXdelta = Matrix(DATA_DIM, N_CLASSES); // DATA_DIM, CHUNK_SIZE * CHUNK_SIZE, N_CLASSES
-
-        Matrix delta_sums = Matrix(1, N_CLASSES); // like b
-
-
-    Matrix lr_grad_W = Matrix(DATA_DIM, N_CLASSES); // like W
[…]

### 2017-02-08 main.cpp
Made cost, acc optional; comment cmake

@@ -63,12 +63,8 @@ int main(int argc, const char* argv[]) {
     Matrix images(n_samples, 784), labels(n_samples, 10);
     read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/test.csv"), images, labels, false);
     NeuralNetwork net;
-    vector<double> cost_history, acc_history;
     cout << setprecision(3);
-    net.train(images, labels, cost_history, acc_history, 100, 200, 0.5);
-    for(auto acc : acc_history) {
-        cout << acc << endl;
-    }
+    auto histories = net.train(images, labels, 100, 200, 0.5, true, true, true);
 //    for (int i = 0; i < acc_history.size(); ++i) {
 //        cout << acc_history[i] << endl;
 //    }

### 2017-02-08 nn.cpp

@@ -6,8 +6,9 @@
 #include "matrix.cpp"
 #include "evaluate.cpp"
 #include <vector>
-#include <chrono>
+#include <chrono> // stopwatch
 #include "mpi.h"
+#include <utility> // pair
 
 
 using namespace std;
@@ -107,12 +108,13 @@ public:
         return cross_entropy(chunk_Y, cY_prob); // cost
     }
 
-    void train(const Matrix& X, const Matrix& Y,
+    pair<vector<double>, vector<double>>
+    train(const Matrix& X, const Matrix& Y,
 //               vector<double>& cost_history=NONE, vector<double>& accuracy_history=NONE,
-               vector<double>& cost_history, vector<double>& accuracy_history,
                const int n_epochs=100, const int batch_size=200, const double lr=0.1,
+                bool compute_cost=true, bool compute_acc=true,
                const bool verbose=true) {
-        bool compute_cost = (cost_history == NONE), compute_acc = (accuracy_history == NONE);
+
         const int n_samples = X.n_rows;
         const int data_dim = X.n_cols;
         const int n_classes = Y.n_cols;
@@ -141,7 +143,8 @@ public:
         Matrix delta_sums(b.n_rows, b.n_cols);
 
         chrono::time_point<chrono::system_clock> start_time, end_time;
-        double total_cost, partial_cost, acc;
+        double acc, total_cost, partial_cost;
+        vector<double> accuracy_history, cost_history;
         vector<int> Y_labels = labels_from_one_hot(Y); // {5, 2, 9, ... }
         int start_index;
 
@@ -225,6 +228,8 @@ public:
 
         // TODO: print total time and average epoch time, max accuracy
         MPI_Finalize();
+
+        return make_pair(accuracy_history, cost_history);
     }
 
     vector<int> predict(const Matrix& X) {

### 2017-02-08 main.cpp
cleanup

@@ -65,9 +65,6 @@ int main(int argc, const char* argv[]) {
     NeuralNetwork net;
     cout << setprecision(3);
     auto histories = net.train(images, labels, 100, 200, 0.5, true, true, true);
-//    for (int i = 0; i < acc_history.size(); ++i) {
-//        cout << acc_history[i] << endl;
-//    }
 
 
 //    Matrix m(4, 2);

### 2017-02-08 nn.cpp

@@ -110,10 +110,8 @@ public:
 
     pair<vector<double>, vector<double>>
     train(const Matrix& X, const Matrix& Y,
-//               vector<double>& cost_history=NONE, vector<double>& accuracy_history=NONE,
-               const int n_epochs=100, const int batch_size=200, const double lr=0.1,
-                bool compute_cost=true, bool compute_acc=true,
-               const bool verbose=true) {
+          const int n_epochs=100, const int batch_size=200, const double lr=0.1,
+          bool compute_cost=true, bool compute_acc=true, bool verbose=true) {
 
         const int n_samples = X.n_rows;
         const int data_dim = X.n_cols;
@@ -126,6 +124,7 @@ public:
         MPI_Comm_size(MPI_COMM_WORLD, &n_processes);
         const int n_workers = n_processes - 1; // how many processes compute gradient in parallel, first is master
         const int chunk_size = batch_size / n_workers; // how many rows each worker gets
+        if (rank != MASTER) verbose = false; // only master is allowed to talk
 
         if (n_samples % chunk_size != 0) throw runtime_error("samples not divisible by chunk");
 
@@ -152,7 +151,7 @@ public:
         random_init(W);
 
         for (int epoch = 1; epoch <= n_epochs; ++epoch) {
-            if (rank == MASTER) {
+            if (verbose) {
                 cout << string_format("Epoch %d / %d", epoch, n_epochs) << flush;
                 start_time = chrono::system_clock::now();
             }
@@ -163,34 +162,32 @@ public:
                     // clear summed from previous batch
                     total_grad_W.clear();
                     total_grad_b.clear();
-                    total_cost = 0;
+                    if (compute_cost) total_cost = 0;
 
                     // Get partial gradients from each worker
                     for (int worker = 1; worker <= n_workers; ++worker) {
                         MPI_Recv(partial_grad_W.data[0], partial_grad_W.n_elements, MPI_DOUBLE, worker, GRAD_W_TAG, MPI_COMM_WORLD, MPI_STATUS_IGNORE);
                         MPI_Recv(partial_grad_b.data[0], partial_grad_b.n_elements, MPI_DOUBLE, worker, GRAD_B_TAG, MPI_COMM_WORLD, MPI_STATUS_IGNORE);
[…]

### 2017-02-08 doc/Paper.tex
Addproblem statement

@@
 \newpage
-\TODO{introduction?}
+\section{Introduction}
+\TODO{more here}
 
@@
 \label{sec:problem}
-Given a picture of a handwritten digit, label it accordingly 0 - 9.
-\TODO{more here}
+Pentru o retea neurala care foloseste un algoritm de antrenare bazat pe metoda gradientului descendent si dispune de o cantitate mare de date de antrenare timpul ca reteaua sa ajunga la rezultate satisfacatoare poate dura chiar si zile pe o implementare secventiala
+Ne propunem sa elaboram o implementare paralela a metodei gradientului descendent.
+Vom folosi pentru validarea implementarii o problema celebra in computer vision, cea a detectarii cifrelor scrise de mana.
+
+For a neural network witch use for training the gradient descendent algorithm and dispose a large amount of training data, the time to reach convergence it can take days in a sequentially implementation.
+We purpose to elaborate a parallel implementation of gradient descendent, and for validate that we use a common problem in computer vision: given a picture of a handwritten digit, label it accordingly 0 - 9.
 
@@
 
-We use a back-propagation neural network:
-\begin{align*}
-W^{(k)} &= W^{(k-1)} + \alpha \nabla_W^{(k-1)} \\ 
-b^{(k)} &= b^{(k-1)} + \alpha \nabla_b^{(k-1)}
-\end{align*}
-Where $W$ is the weights matrix, $10 \times 784$ -- the number of classes( one for each digit) and the dimension of an image ($28 \times 28$).
+We use a feed-forward neural network with an one fully connected layer, which predict the digit that has been in input image.
+For an image $x$, initialy $28 \times 28$, we flatten this array into a vector of $784 = 28 \times 28$.
+
+\begin{align}
+y = softmax(Wx + b) \\
+image\_label &= argmax(y)
+\end{align}
+
+In $(1)$, $W$ is the weights matrix, $10 \times 784$ -- the number of classes(one for each digit) and the dimension of an image ($784$).
 $b$ is the bias vector, $1 \times 10$, corresponding to the free element.
-$W^{(k)}$ denotes the weights at step $k$.
-$\nabla$ represents the update which must produce to improve prediction.
 \\
 
-The activation function is softmax:
-$$f(x_i)=\frac{exp(x_i)}{\sum_j exp(x_j)}$$
+The activation function output a vector of probabilities:
+$$softmax(x_i)=\frac{exp(x_i)}{\sum_j exp(x_j)}$$
 Where $exp(x)$ represents $e^x$, Euler's constant. It \textit{normalizes} the vector, causing the greatest elements to have logarithmic relative value.
 \\
+
+$(2)$ With $y$ vector of probabilities, $argmax$ take the index of the maximum value and results the prediction label.
 
@@
 \paragraph{Training} We employ gradient-descent on mini-batches. 
-Early-stopping prevents the model from further learning when performance on the validation split starts decreasing.
-Random sampling averts cycles when iterating through the dataset.
-
-\paragraph{Prediction}
-\TODO{explain prediction?}
+\begin{align*}
+W^{(k)} &= W^{(k-1)} + \alpha \nabla_W^{(k-1)} \\ 
+b^{(k)} &= b^{(k-1)} + \alpha \nabla_b^{(k-1)}
+\end{align*}
+Where $W$ is the weights matrix, $10 \times 784$ -- the number of classes( one for each digit) and the dimension of an image ($28 \times 28$).
+$b$ is the bias vector, $1 \times 10$, corresponding to the free element.
+$W^{(k)}$ denotes the weights at step $k$.
+$\nabla$ represents the update which must produce to improve prediction.
+\\
+Early-stopping prevents the model from further learning when performance on the validation split starts decreasing. Random sampling averts cycles when iterating through the dataset.
 

### 2017-02-08 main.cpp
dynamic W & b

@@ -53,32 +53,11 @@ int main(int argc, const char* argv[]) {
 //        return 0;
 //    }
 
-//    double* v = new double[5];
-//    for (int i = 0; i < 5; ++i)
-//        v[i] = 100+i;
-//    double* w = v+2;
-//    cout << w[0];
-
-    int n_samples = 2000;
-    Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv(string("/Users/Stefan/Projects/mnist-in-mpi/data/test.csv"), images, labels, false);
-    NeuralNetwork net;
     cout << setprecision(3);
+
+    int n_samples = 2000; Matrix images(n_samples, 784), labels(n_samples, 10);
+    read_from_csv("data/test.csv", images, labels, false);
+    NeuralNetwork net(784, 10);
     auto histories = net.train(images, labels, 100, 200, 0.5, true, true, true);
-
-
-//    Matrix m(4, 2);
-//    int k = 0;
-//    for (int i = 0; i < m.n_rows; ++i)
-//        for (int j = 0; j < m.n_cols; ++j)
-//            m.data[i][j] = k++;
-//    cout << m;
-//
-//    Matrix chunk(2, 2);
-//    take_chunk(m, 1, chunk);
-//    cout << endl << chunk;
-
-
-    return 0;
 }
 

### 2017-02-08 nn.cpp

@@ -66,21 +66,16 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 
 class NeuralNetwork {
 
-    const int N_CLASSES = 10; // 10 digits from 0 to 9
-    const int DATA_DIM = 784; // 28*28 = 784 pixels for an image
-
-
-    // X dimensions: N_SAMPLES x DATA_DIM
-    // Y dimensions: N_SAMPLES x N_CLASSES
-
-    Matrix W = Matrix(DATA_DIM, N_CLASSES);        // TODO? more layers
-    Matrix b = Matrix(1, N_CLASSES);
-
-
+    Matrix W;        // TODO? more layers
+    Matrix b;
 
 public:
+    const int N_CLASSES;
+    const int DATA_DIM;
 
-    NeuralNetwork() { }
+    NeuralNetwork(const int data_dim, const int n_classes) :
+            N_CLASSES(n_classes), DATA_DIM (data_dim),
+            W(data_dim, n_classes), b(1, n_classes) { }
 
 //    NeuralNetwork(string path) {
 //        this->load(path);
@@ -104,14 +99,13 @@ public:
         col_wise_sums(delta, delta_sums);
         scalar_mult(contribution, delta_sums, partial_grad_b); // grad_b = 1/n * col_wise_sums(delta)
 
-        // TODO maybe
-        return cross_entropy(chunk_Y, cY_prob); // cost
+        return compute_cost ? cross_entropy(chunk_Y, cY_prob) : 0;
     }
 
     pair<vector<double>, vector<double>>
     train(const Matrix& X, const Matrix& Y,
           const int n_epochs=100, const int batch_size=200, const double lr=0.1,
-          bool compute_cost=true, bool compute_acc=true, bool verbose=true) {
+          bool compute_acc=true, bool compute_cost=true, bool verbose=true) {
 
         const int n_samples = X.n_rows;
         const int data_dim = X.n_cols;
@@ -197,7 +191,7 @@ public:
                     // Send partial gradients to master
                     MPI_Send(partial_grad_W.data[0], partial_grad_W.n_elements, MPI_DOUBLE, MASTER, GRAD_W_TAG, MPI_COMM_WORLD);
                     MPI_Send(partial_grad_b.data[0], partial_grad_b.n_elements, MPI_DOUBLE, MASTER, GRAD_B_TAG, MPI_COMM_WORLD);
[…]

### 2017-02-08 main.cpp
Redo train_and_save

@@ -1,63 +1,62 @@
 #include <iostream>
-#include "reading.cpp"
-#include "nn.cpp"
-#include "matrix.cpp"
 #include <iomanip>
+#include "mnist_io.cpp"
+#include "nn.cpp"
 
 using namespace std;
 
-//void train_and_save() {
-//    // TODO: names as additional parameters
-//    Matrix images, labels;
-//    read_data("data/train.csv", images, labels);
-//
-//    NeuralNetwork net;
-//    vector<double> costs, accs;
-//    net.train(images, labels, costs, accs, 35, 100, .1);
-//
-//    net.save("models/trained.nn");
-//    ofstream f("histories/training.txt");
-//    f << costs << endl << accs;
-//    f.close();
-//}
 
-//int predict(string pixels_string) {
-//    // TODO: names as additional parameters
-//    vector<double> pixels = pixels_from_string(pixels_string);
-//    NeuralNetwork net("models/trained_a.nn");
-//
-//    double confidence;
-//    int digit_predicted = net.predict_one(pixels, confidence);
-//    cout << digit_predicted << endl << confidence << endl;
-//
-//    return digit_predicted;
-//}
-
-int main(int argc, const char* argv[]) {
-//    if (argc < 2) {
-//        cout << "usage: mnist_in_mpi.out (train|predict)" << endl;
-//        return 1;
-//    }
-//
-//    if (string(argv[1]) == "train") {
-//        train_and_save();
-//        return 0;
-//    }
-//
-//    if (string(argv[1]) == "predict") {
-//        if (argc < 3) {
-//            cout << "usage mnist_in_mpi.out predict <pixels>" << endl;
-//            return 1;
-//        }
-//        predict(argv[2]);
-//        return 0;
-//    }
-
+void train_and_save() {
     cout << setprecision(3);
 
-    int n_samples = 2000; Matrix images(n_samples, 784), labels(n_samples, 10);
-    read_from_csv("data/test.csv", images, labels, false);
-    NeuralNetwork net(784, 10);
-    auto histories = net.train(images, labels, 100, 200, 0.5, true, true, true);
+    // Read
+    int n_samples = 1000; Matrix images(n_samples, IMAGE_PIXELS), labels(n_samples, N_DIGITS);
+    read_from_csv("data/test.csv", images, labels, false);    // TODO: names & params as additional args
+
+    // Train
+    NeuralNetwork net(IMAGE_PIXELS, N_DIGITS);
[…]

### 2017-02-08 matrix.cpp

@@ -51,6 +51,7 @@ std::ostream& operator<<(std::ostream& os, const Matrix& matrix) {
     return os;
 }
 std::istream& operator>>(std::istream& is, Matrix& matrix) {
+    cout << "reading " << matrix.n_rows << " x " << matrix.n_cols << endl;
     for (int i = 0; i < matrix.n_rows; ++i)
         for (int j = 0; j < matrix.n_cols; ++j)
             is >> matrix.data[i][j];

### 2017-02-08 mnist_io.cpp

@@ -9,8 +9,12 @@
 
 using namespace std;
 
+
 const int PIXEL_DIM = 255;
 const int IMAGE_SIDE = 28;
+const int IMAGE_PIXELS = IMAGE_SIDE * IMAGE_SIDE;
+const int N_DIGITS = 10; // hah
+
 
 void read_from_csv(string filename, Matrix& images, Matrix& labels, bool verbose=true) {
     // images.n_rows indicates how many images to read from the csv
@@ -58,3 +62,10 @@ void print_image(const Matrix& images, const int image_n) {
     }
     cout << endl;
 }
+
+void parse_image(const string& str, Matrix& X) {
+    istringstream stream(str);
+    for (int i = 0; i < X.n_cols; i++)
+        stream >> X.data[0][i]; // on first row
+}
+

### 2017-02-08 nn.cpp

@@ -16,9 +16,9 @@ using namespace std;
 
 /*** MPI constants ***/
 const int MASTER = 0;
-const int GRAD_W_TAG    = 0;
-const int GRAD_B_TAG    = 1;
-const int COST_TAG      = 2;
+const int GRAD_W_TAG = 0;
+const int GRAD_B_TAG = 1;
+const int COST_TAG   = 2;
 
 static vector<double> NONE = vector<double>();
 
@@ -66,10 +66,12 @@ void random_init(Matrix& W, double mean=0., double std=1.) {
 
 class NeuralNetwork {
 
-    Matrix W;        // TODO? more layers
-    Matrix b;
+private:
 
 public:
+    Matrix b;
+    Matrix W;        // TODO? more layers
+
     const int N_CLASSES;
     const int DATA_DIM;
 
@@ -77,9 +79,12 @@ public:
             N_CLASSES(n_classes), DATA_DIM (data_dim),
             W(data_dim, n_classes), b(1, n_classes) { }
 
-//    NeuralNetwork(string path) {
-//        this->load(path);
-//    }
+    NeuralNetwork(string path) :
+        N_CLASSES(10), DATA_DIM(784),
+        W(784, 10), b(1, 10) // TODO
+    {
+        this->load(path);
+    }
 
     double grad(const Matrix& chunk_X, const Matrix& chunk_Y,
                 Matrix& partial_grad_W, Matrix& partial_grad_b,
@@ -144,9 +149,14 @@ public:
         // Initial weights
         random_init(W);
 
+        if (verbose)
+            cout << "Training on " << n_samples / 1000 << "k samples"
+                 << " with " << n_workers << " workers"
+                 << " (" << chunk_size << " chunk size)" << endl;
+
         for (int epoch = 1; epoch <= n_epochs; ++epoch) {
             if (verbose) {
-                cout << string_format("Epoch %d / %d", epoch, n_epochs) << flush;
+                cout << string_format("Epoch %d/%d", epoch, n_epochs) << flush;
                 start_time = chrono::system_clock::now();
             }
 
@@ -253,15 +263,15 @@ public:
         file.close();
     }
 
-//    void load(string path) {
-//        ifstream file(path);
-//        if (!file.is_open())
-//            throw runtime_error(string_format("Could not open for model load: " + path));
-//
-//        file >> *this;
-//
-//        file.close();
-//    }
+    void load(string path) {
+        ifstream file(path);
+        if (!file.is_open())
[…]

### 2017-02-08 utils.cpp

@@ -28,38 +28,15 @@ std::string string_format(const std::string fmt_str, ...) {
     return std::string(formatted.get());
 }
 
-//template <class T>
-//ostream& operator<<(ostream& os, const vector<T>& v) {
-//    for (const auto& x : v)
-//        os << x << ' ';
-//    return os;
-//}
-//
-//template <class T>
-//istream& operator>>(istream& is, vector<T>& v) {
-//    string line;
-//    getline(is, line);
-//    istringstream ss(line);
-//
-//    T x;
-//    while (ss >> x) // until end of line
-//        v.push_back(x);
-//
-//    return is;
-//}
-//
-//vector<double> pixels_from_string(const string& s) {
-//    vector<double> v;
-//    istringstream ss(s);
-//    double x;
-//    while (ss >> x) {
-//        v.push_back(x);
-//        if (ss.peek() == ',')
-//            ss.ignore();
-//    }
-//    return v;
-//}
+void save_csv(pair<vector<double>,vector<double>> histories, string path) {
+    auto accs = histories.first, costs = histories.second;
 
+    ofstream f(path);
+    f << "epoch,accuracy,cost" << endl; // header
+    for (int i = 0; i < accs.size(); ++i)
+        f << i+1 << accs[i] << costs[i] << endl;
+    f.close();
+}
 
 // for 2 (and 5 classes) ~> {0 0 1 0 0}
 vector<double> make_one_hot(int value, int n_classes=10) {

### 2017-02-08 main.cpp
Rewrite in progress: predict_one

@@ -31,13 +31,15 @@ void predict(string pixels_str) {
     NeuralNetwork net("models/trained.nn");     // TODO: names as additional args
 
     // Predict
-    double confidence;
-//    int digit_predicted = net.predict_one(pixels, confidence);
-//    cout << digit_predicted << endl << confidence << endl;
+    auto prediction = net.predict_one(input);
+    cout << endl << "prediction: ";
+    cout << prediction.first << endl << prediction.second << endl;
 }
 
 int main(int argc, const char* argv[]) {
-    train_and_save();
+    cout << setprecision(2);
+    string dummy = "0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.035681610247026534,0.14468696902365705,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.7847340216965103,0.9996078943928898,0.3182590511044308,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.9332113449222323,0.9998692981309633,0.46660567246111617,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.9332113449222323,0.9998692981309633,0.46660567246111617,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.9332113449222323,0.9998692981309633,0.46660567246111617,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.9465429355639786,0.9998692981309633,0.4667363743301529,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.10351588027708797,0.9998692981309633,0.9998692981309633,0.3609985622794406,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.4506600444386355,0.9998692981309633,0.9998692981309633,0.1388053849170043,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.8065612338256437,0.9998692981309633,0.80708
[…]

### 2017-02-08 matrix.cpp

@@ -51,7 +51,6 @@ std::ostream& operator<<(std::ostream& os, const Matrix& matrix) {
     return os;
 }
 std::istream& operator>>(std::istream& is, Matrix& matrix) {
-    cout << "reading " << matrix.n_rows << " x " << matrix.n_cols << endl;
     for (int i = 0; i < matrix.n_rows; ++i)
         for (int j = 0; j < matrix.n_cols; ++j)
             is >> matrix.data[i][j];

### 2017-02-08 mnist_io.cpp

@@ -63,9 +63,12 @@ void print_image(const Matrix& images, const int image_n) {
     cout << endl;
 }
 
-void parse_image(const string& str, Matrix& X) {
+void parse_image(string str, Matrix& X) {
+    for (int i = 0; i < str.size(); ++i)
+        if (str[i] == ',') str[i] = ' ';
     istringstream stream(str);
     for (int i = 0; i < X.n_cols; i++)
         stream >> X.data[0][i]; // on first row
+    cout << X.data[0][0]<<endl;
 }
 

### 2017-02-08 nn.cpp

@@ -122,6 +122,8 @@ public:
         MPI_Comm_rank(MPI_COMM_WORLD, &rank);
         MPI_Comm_size(MPI_COMM_WORLD, &n_processes);
         const int n_workers = n_processes - 1; // how many processes compute gradient in parallel, first is master
+        if (n_workers < 1) throw runtime_error("no workers!");
+
         const int chunk_size = batch_size / n_workers; // how many rows each worker gets
         if (rank != MASTER) verbose = false; // only master is allowed to talk
 
@@ -240,16 +242,24 @@ public:
         softmax(XWb, Y_prob); // Y_prob = softmax(X * W + b)
         return argmax(Y_prob);
     }
-//
-//    int predict_one(const vector<double>& pixels, double& confidence) {
-//        // TODO: refactor to use .predict()
-//        Matrix X = {pixels};
-//        vector<double> y_prob = softmax(X * W + b)[0]; // just one image, take the first row
-//
-//        int digit_predicted = argmax(y_prob);
-//        confidence = y_prob[digit_predicted];
-//        return digit_predicted;
-//    }
+
+    pair<int, double> predict_one(const Matrix& X) {
+        // TODO: refactor to use NeuralNetwork::predict
+        const int n_samples = X.n_rows, n_classes = b.n_cols;
+        Matrix XW = Matrix(n_samples, n_classes), XWb(n_samples, n_classes), Y_prob(n_samples, n_classes);
+
+        cout << X;
+        cout << W;
+        cout << XW;
+
+        dot(X, W, XW); // XW = X * W
+        add_to_each(XW, b, XWb); // XWb = X * W + b
+        softmax(XWb, Y_prob); // Y_prob = softmax(X * W + b)
+
+        int digit_predicted = argmax(Y_prob)[0]; // first image, the only one
+        double confidence = Y_prob.data[0][digit_predicted];
+        return make_pair(digit_predicted, confidence);
+    }
 
 
     /*** Serialization ***/

### 2017-02-08 server.py

@@ -1,9 +1,7 @@
 #! /usr/bin/env python3
-
 import os
 from flask import Flask, render_template, jsonify, request
 app = Flask(__name__, template_folder='frontend', static_folder='frontend')
-
 
 @app.route('/')
 def index():
@@ -12,7 +10,7 @@ def index():
 @app.route('/recognize')
 def compute():
     pixels = request.args['pixels']
-    command = './mnist_in_mpi.out predict ' + pixels
+    command = '../mnist_in_mpi.out predict ' + pixels
 
     output = os.popen(command).read()  # very hackish, very bad, don't do this
     digit_predicted = int(output.split()[0])

### 2017-02-08 utils.cpp

@@ -45,3 +45,11 @@ vector<double> make_one_hot(int value, int n_classes=10) {
     return result;
 }
 
+template <class T>
+ostream& operator<<(ostream& os, const vector<T>& v) {
+    os << '[';
+    for (const auto& x : v)
+        os << x << ' ';
+    return os << ']';
+}
+

### 2017-02-08 evaluate.cpp
Fix predict_one

@@ -31,8 +31,7 @@ vector<int> labels_from_one_hot(const Matrix& matrix) {
 
 vector<int> argmax(const Matrix& matrix) {
     vector<int> result(matrix.n_rows, 0);
-    double row_max;
-    int col_of_max;
+    int col_of_max = 0;
 
     for (int row = 0; row < matrix.n_rows; ++row) {
         double row_max = matrix.data[row][0];

### 2017-02-08 main.cpp

@@ -32,15 +32,12 @@ void predict(string pixels_str) {
 
     // Predict
     auto prediction = net.predict_one(input);
-    cout << endl << "prediction: ";
     cout << prediction.first << endl << prediction.second << endl;
 }
 
 int main(int argc, const char* argv[]) {
-    cout << setprecision(2);
-    string dummy = "0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.035681610247026534,0.14468696902365705,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.7847340216965103,0.9996078943928898,0.3182590511044308,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.9332113449222323,0.9998692981309633,0.46660567246111617,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.9332113449222323,0.9998692981309633,0.46660567246111617,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.9332113449222323,0.9998692981309633,0.46660567246111617,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.9465429355639786,0.9998692981309633,0.4667363743301529,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.10351588027708797,0.9998692981309633,0.9998692981309633,0.3609985622794406,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.4506600444386355,0.9998692981309633,0.9998692981309633,0.1388053849170043,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.8065612338256437,0.9998692981309633,0.8070840413017906,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.18664226898444647,1,0.9998692981309633,0.4289635341785388,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0
[…]

### 2017-02-08 mnist_io.cpp

@@ -69,6 +69,5 @@ void parse_image(string str, Matrix& X) {
     istringstream stream(str);
     for (int i = 0; i < X.n_cols; i++)
         stream >> X.data[0][i]; // on first row
-    cout << X.data[0][0]<<endl;
 }
 

### 2017-02-08 nn.cpp

@@ -248,10 +248,6 @@ public:
         const int n_samples = X.n_rows, n_classes = b.n_cols;
         Matrix XW = Matrix(n_samples, n_classes), XWb(n_samples, n_classes), Y_prob(n_samples, n_classes);
 
-        cout << X;
-        cout << W;
-        cout << XW;
-
         dot(X, W, XW); // XW = X * W
         add_to_each(XW, b, XWb); // XWb = X * W + b
         softmax(XWb, Y_prob); // Y_prob = softmax(X * W + b)

### 2017-02-08 nn.cpp
Add shuffle & anneal_every params to train(), run benchmarks in Paper

@@ -111,13 +111,19 @@ public:
 
     pair<vector<double>, vector<double>>
     train(const Matrix& X, const Matrix& Y,
-          const int n_epochs=100, const int batch_size=200, const double lr=0.1,
-          bool compute_acc=true, bool compute_cost=true, bool verbose=true) {
+          const int n_epochs=100, const int batch_size=200,
+          const double lr=0.1, int anneal_every=-1,
+          const bool shuffle=true, // TODO regularization parameter
+          const bool compute_acc=true, const bool compute_cost=true, bool verbose=true) {
 
         const int n_samples = X.n_rows;
         const int data_dim = X.n_cols;
         const int n_classes = Y.n_cols;
         double lr_anneal = lr;
+
+        if (anneal_every == -1)
+            anneal_every = n_epochs / 3; // by default, anneal after every third of n_epochs
+
         // MPI
         int rank, n_processes;
         MPI_Init(NULL, NULL);
@@ -164,7 +170,7 @@ public:
                 start_time = chrono::system_clock::now();
             }
             // learning rate anneal
-            if (epoch % (n_epochs/3) == 0)
+            if (epoch % anneal_every == 0)
                 lr_anneal /= 2.;
 
             for (int batch_start = 0; batch_start < X.n_rows; batch_start += batch_size) {
@@ -187,7 +193,7 @@ public:
                         if (compute_cost) total_cost += partial_cost;
                     }
 
-                    //After receive
+                    // After receive
                     scalar_mult(lr_anneal, total_grad_W, lr_grad_W);                     // TODO? regularization
                     scalar_mult(lr_anneal, total_grad_b, lr_grad_b);
                     sub_from(W, lr_grad_W); // W = W - lr * grad_W
@@ -198,7 +204,7 @@ public:
 
 
                 if (rank != MASTER) { // worker
-                     start_index = batch_start + (rank - 1) * chunk_size;                    // TODO? make random batch generator
+                    start_index = batch_start + (rank - 1) * chunk_size;
                     start_index = rand_int(0, n_samples - chunk_size - 1);
                     take_chunk(X, start_index, chunk_X); // chunk_X = X[batch_start ... batch_start + CHUNK_SIZE]
                     take_chunk(Y, start_index, chunk_Y);

### 2017-02-09 main.cpp
Add regularization

@@ -7,19 +7,32 @@ using namespace std;
 
 
 void train_and_save() {
+    int rank, n_processes;
+    MPI_Init(NULL, NULL);
+    MPI_Comm_rank(MPI_COMM_WORLD, &rank);
+    MPI_Comm_size(MPI_COMM_WORLD, &n_processes);
+
     cout << setprecision(3);
 
-    // Read
-    int n_samples = 10000; Matrix images(n_samples, IMAGE_PIXELS), labels(n_samples, N_DIGITS);
-    read_from_csv("data/test.csv", images, labels, false);    // TODO: names & params as additional args
+    // Read Train
+    int n_samples = 60000; Matrix images(n_samples, IMAGE_PIXELS), labels(n_samples, N_DIGITS);
+    read_from_csv("data/train.csv", images, labels, false);    // TODO: names & params as additional args
+
+    // Read Test
+    int n_samples_test = 10000; Matrix images_test(n_samples_test, IMAGE_PIXELS), labels_test(n_samples_test, N_DIGITS);
+    read_from_csv("data/test.csv", images_test, labels_test, false);    // TODO: names & params as additional args
 
     // Train
     NeuralNetwork net(IMAGE_PIXELS, N_DIGITS);
-    auto histories = net.train(images, labels, 50, 200, 1.);
+    auto histories = net.train(images, labels, images_test, labels_test, 100, 200, .5, true, true);
+    if (rank == 0) {
+        // Save
+        net.save("models/trained.nn");
+        save_csv(histories, "histories/training.csv");
+    }
 
-    // Save
-    net.save("models/trained.nn");
-    save_csv(histories, "histories/training.csv");
+    MPI_Finalize();
+
 }
 
 void predict(string pixels_str) {
@@ -36,11 +49,11 @@ void predict(string pixels_str) {
 }
 
 int main(int argc, const char* argv[]) {
[…]

### 2017-02-09 nn.cpp

@@ -9,8 +9,9 @@
 #include <chrono> // stopwatch
 #include "mpi.h"
 #include <utility> // pair
-#include <stdlib.h>
-
+#include <stdlib.h> // rand()
+#include <tuple>        // std::tuple, std::get, std::tie, std::ignore
+#include <time.h>       /* time */
 
 using namespace std;
 
@@ -97,30 +98,32 @@ public:
         softmax(cXWb, cY_prob); // cY_prob = softmax(X * W + b)
         sub(cY_prob, chunk_Y, delta); // delta = cY_prob - Y
 
-        double contribution = 1/double(chunk_X.n_rows);
+        // double contribution = 1/double(batch_size);
 
         transpose(chunk_X, trX); // trX = transpose(X)
-        dot(trX, delta, trXdelta); // trXdelta = transpose(X) * delta
-        scalar_mult(contribution, trXdelta, partial_grad_W); // grad_W = 1/n * transpose(chunk_X) * delta
+        dot(trX, delta, partial_grad_W); // trXdelta = transpose(X) * delta
+        // scalar_mult(contribution, trXdelta, partial_grad_W); // grad_W = 1/n * transpose(chunk_X) * delta
 
-        col_wise_sums(delta, delta_sums);
-        scalar_mult(contribution, delta_sums, partial_grad_b); // grad_b = 1/n * col_wise_sums(delta)
+        col_wise_sums(delta, partial_grad_b);
+        // scalar_mult(contribution, delta_sums, partial_grad_b); // grad_b = 1/n * col_wise_sums(delta)
 
         return compute_cost ? cross_entropy(chunk_Y, cY_prob) : 0;
     }
 
-    pair<vector<double>, vector<double>>
+    tuple<vector<double>, vector<double>, vector<double>>
     train(const Matrix& X, const Matrix& Y,
+          const Matrix& X_test, const Matrix& Y_test,
           const int n_epochs=100, const int batch_size=200, const double lr=0.1,
           bool compute_acc=true, bool compute_cost=true, bool verbose=true) {
 
+        const double lambda_reg = 5.0;
         const int n_samples = X.n_rows;
         const int data_dim = X.n_cols;
         const int n_classes = Y.n_cols;
         double lr_anneal = lr;
         // MPI
         int rank, n_processes;
-        MPI_Init(NULL, NULL);
+        // MPI_Init(NULL, NULL);
         MPI_Comm_rank(MPI_COMM_WORLD, &rank);
[…]

### 2017-02-09 utils.cpp

@@ -4,9 +4,18 @@
 #include <string.h>
 #include <stdarg.h>  // For va_start, etc.
 #include <memory>    // For std::unique_ptr
+#include <tuple>        // std::tuple, std::get, std::tie, std::ignore
+
 
 using namespace std;
 
+template <class T>
+ostream& operator<<(ostream& os, const vector<T>& v) {
+    os << '[';
+    for (const auto& x : v)
+        os << x << ' ';
+    return os << ']';
+}
 
 // source: http://stackoverflow.com/questions/2342162/stdstring-formatting-like-sprintf
 std::string string_format(const std::string fmt_str, ...) {
@@ -28,13 +37,13 @@ std::string string_format(const std::string fmt_str, ...) {
     return std::string(formatted.get());
 }
 
-void save_csv(pair<vector<double>,vector<double>> histories, string path) {
-    auto accs = histories.first, costs = histories.second;
-
+void save_csv(tuple<vector<double>,vector<double>, vector<double>> histories, string path) {
+    vector<double> acc_train, acc_test, costs;
+    tie (acc_train, acc_test, costs) = histories;
     ofstream f(path);
-    f << "epoch,accuracy,cost" << endl; // header
-    for (int i = 0; i < accs.size(); ++i)
-        f << i+1 << accs[i] << costs[i] << endl;
+    f << "epoch,accuracy_train,accuracy_test,cost" << endl; // header
+    for (int i = 0; i < acc_train.size(); ++i)
+        f << i+1 << "," << acc_train[i] << "," << acc_test[i] << "," << costs[i] << endl;
     f.close();
 }
 
@@ -44,12 +53,3 @@ vector<double> make_one_hot(int value, int n_classes=10) {
     result[value] = 1;
     return result;
 }
-
-template <class T>
-ostream& operator<<(ostream& os, const vector<T>& v) {
-    os << '[';
-    for (const auto& x : v)
-        os << x << ' ';
-    return os << ']';
-}
-

### 2017-02-09 doc/Paper.tex

@@
 \section{Introduction}
-\TODO{more here}
+Proiectul are ca scop evidentierea rolului paralelismului in programare.
+Avand vedere cantitatea mare de date cu care se confrunta aplicatiile din ziua de astazi, este nevoie ca acestea sa ruleze anumite sarcini concomitent.
+Pentru a sublinia imbunatatirile aduse de paralelism, s-a implementat in MPI antrenarea unei retele neuronale.
+In continuare se vor detalia, datele folosite in model, reteaua construita, metodele de paralelism, rezultatele, dar si interfata.
 
@@
 \begin{align}
-y = softmax(Wx + b) \\
-image\_label &= argmax(y)
+y &= softmax(Wx + b) \\
+label &= argmax(y)
 \end{align}
@@
   ylabel=Seconds,
-  bar width=.8cm,
+  bar width=1cm,
+  enlarge x limits=0.2,
+%   enlarge y limits=0.18,
   axis line style={black!65, line width=.1pt}, % thinner border
-  symbolic x coords={1,2,4,6,8,12,16}, % equidistant bars
+  symbolic x coords={1,2,4,8,16}, % equidistant bars
   every x tick/.style={white}, % hide x ticks (make them white)
@@
   ymajorgrids=true, % show horizontal grid lines
-  grid style={line width=.1pt, draw=gray!10}, % gray grid lines
+  grid style={line width=.1pt, draw=gray!20}, % gray grid lines
   ]
   \addplot[ybar, black!40, fill=black!20] coordinates {
-      (1,8) 
-      (2,5)
-      (4,4) 
-      (6,3) 
-      (8,2.5) 
-      (12,2) 
-      (16,1.5)    
+      (1,5.27) 
+      (2,2.81)
+      (4,1.51)
+      (8,0.78)
+      (16,0.79)
       };
@@
 
-\TODO{comment on chart}
+\noindent The benchmarks were done on the MNIST training set of 60 000 images with a batch size of 1000, on a machine with 20 processors @ 2.60GHz.
+
+The training time is almost halved when going from one (5.27s) to two (2.81s) processors: 88\% improvement.
+Further 86\% improvement is achieved on four processors (1.51s), less than a third (29\%) of the single-process time.
+Doubling the number of processes to eight yields a further 94\% improvement (0.78s), making it only 15\% of the single-process time.
+This is notable when considering the perfect improvement possible of 12.5\%.
+
+For more than eight cores, the communication overhead starts to take its toll.
+No further performance improvement is achieved.
+Chunk sizes get increasingly small (1000, 500, 250, 125, 63), hitting an inflection point at the eight-processes mark.
 

### 2017-02-09 main.cpp
cleanup

@@ -15,16 +15,16 @@ void train_and_save() {
     cout << setprecision(3);
 
     // Read Train
-    int n_samples = 60000; Matrix images(n_samples, IMAGE_PIXELS), labels(n_samples, N_DIGITS);
+    int n_samples = 10000; Matrix images(n_samples, IMAGE_PIXELS), labels(n_samples, N_DIGITS);
     read_from_csv("data/train.csv", images, labels, false);    // TODO: names & params as additional args
 
     // Read Test
-    int n_samples_test = 10000; Matrix images_test(n_samples_test, IMAGE_PIXELS), labels_test(n_samples_test, N_DIGITS);
+    int n_samples_test = 2000; Matrix images_test(n_samples_test, IMAGE_PIXELS), labels_test(n_samples_test, N_DIGITS);
     read_from_csv("data/test.csv", images_test, labels_test, false);    // TODO: names & params as additional args
 
     // Train
     NeuralNetwork net(IMAGE_PIXELS, N_DIGITS);
-    auto histories = net.train(images, labels, images_test, labels_test, 100, 200, .5, true, true);
+    auto histories = net.train(images, labels, images_test, labels_test, 5, 200, .5, true, true);
     if (rank == 0) {
         // Save
         net.save("models/trained.nn");
@@ -32,7 +32,6 @@ void train_and_save() {
     }
 
     MPI_Finalize();
-
 }
 
 void predict(string pixels_str) {
@@ -51,9 +50,9 @@ void predict(string pixels_str) {
 int main(int argc, const char* argv[]) {
[…]

### 2017-02-09 nn.cpp

@@ -83,8 +83,8 @@ public:
             W(data_dim, n_classes), b(1, n_classes) { }
 
     NeuralNetwork(string path) :
-        N_CLASSES(10), DATA_DIM(784),
-        W(784, 10), b(1, 10) // TODO
+            N_CLASSES(10), DATA_DIM(784),
+            W(784, 10), b(1, 10) // TODO
     {
         this->load(path);
     }
@@ -114,22 +114,19 @@ public:
     train(const Matrix& X, const Matrix& Y,
           const Matrix& X_test, const Matrix& Y_test,
           const int n_epochs=100, const int batch_size=200,
-          const double lr=0.1, int anneal_every=-1,
-          const bool shuffle=true, // TODO regularization parameter
+          double lr=0.1, int anneal_every=-1,
+          const bool shuffle=true, const double lambda_reg = 5.0,
           const bool compute_acc=true, const bool compute_cost=true, bool verbose=true) {
 
-        const double lambda_reg = 5.0;
         const int n_samples = X.n_rows;
         const int data_dim = X.n_cols;
         const int n_classes = Y.n_cols;
-        double lr_anneal = lr;
 
         if (anneal_every == -1)
             anneal_every = n_epochs / 3; // by default, anneal after every third of n_epochs
 
         // MPI
         int rank, n_processes;
-        // MPI_Init(NULL, NULL);
         MPI_Comm_rank(MPI_COMM_WORLD, &rank);
         MPI_Comm_size(MPI_COMM_WORLD, &n_processes);
         const int n_workers = n_processes - 1; // how many processes compute gradient in parallel, first is master
@@ -154,15 +151,15 @@ public:
         Matrix delta_sums(b.n_rows, b.n_cols);
 
         chrono::time_point<chrono::system_clock> start_time, end_time;
-        double acc, total_cost, partial_cost;
-        vector<double> accuracy_train, accuracy_test, cost_history;
+        double accuracy_train, accuracy_test, total_cost, partial_cost;
+        vector<double> accuracy_train_history, accuracy_test_history, cost_history;
         vector<int> Y_labels = labels_from_one_hot(Y); // {5, 2, 9, ... }
         vector<int> Y_labels_test = labels_from_one_hot(Y_test); // {5, 2, 9, ... }
 
         int start_index;
-        // seed
-        srand(time(NULL) + rank);
 
+        // seed
+        srand(time(NULL) * rank);
         // Initial weights
         random_init(W);
 
[…]

### 2017-02-09 doc/Paper.tex
update paper, add ss to readme

@@
 
-\usepackage[automark]{scrpage2} % heading
-  \pagestyle{scrheadings}
-  \ihead[]{Handwritten character recognition in MPI} % left header
-  \ohead[]{\today} % right header
-  \cfoot[]{\pagemark} % footer
-  \setheadsepline[122mm]{0.3mm}
+% \usepackage[automark]{scrpage2} % heading
+%   \pagestyle{scrheadings}
+%   \ihead[]{Handwritten character recognition in MPI} % left header
+%   \ohead[]{\today} % right header
+%   \cfoot[]{\pagemark} % footer
+%   \setheadsepline[122mm]{0.3mm}
 
@@
 \section{Introduction}
-Proiectul are ca scop evidentierea rolului paralelismului in programare.
-Avand vedere cantitatea mare de date cu care se confrunta aplicatiile din ziua de astazi, este nevoie ca acestea sa ruleze anumite sarcini concomitent.
-Pentru a sublinia imbunatatirile aduse de paralelism, s-a implementat in MPI antrenarea unei retele neuronale.
-In continuare se vor detalia, datele folosite in model, reteaua construita, metodele de paralelism, rezultatele, dar si interfata.
+The project aims to highlight the role of parallelism in programming, especially in Machine Learning tasks.
+Considering the vast amount of data today's applications have to deal with, it is paramount to run tasks concomitantly.
+To underline  improvements bought by parallelism, we implemented  a neural network training module in MPI.
+
+We will present the details, data used, model architecture, parallelization strategy and an interactive interface.
 
@@
 \label{sec:problem}
-Pentru o retea neurala care foloseste un algoritm de antrenare bazat pe metoda gradientului descendent si dispune de o cantitate mare de date de antrenare timpul ca reteaua sa ajunga la rezultate satisfacatoare poate dura chiar si zile pe o implementare secventiala
-Ne propunem sa elaboram o implementare paralela a metodei gradientului descendent.
-Vom folosi pentru validarea implementarii o problema celebra in computer vision, cea a detectarii cifrelor scrise de mana.
-
-For a neural network witch use for training the gradient descendent algorithm and dispose a large amount of training data, the time to reach convergence it can take days in a sequentially implementation.
-We purpose to elaborate a parallel implementation of gradient descendent, and for validate that we use a common problem in computer vision: given a picture of a handwritten digit, label it accordingly 0 - 9.
+Given a picture of a handwritten digit, label it accordingly 0 - 9.
+This is a classic problem in Computer Vision on which we apply a neural network trained using gradient descent.
+We seek to reach satisfactory model performance and fast training time, using parallelism.
 
@@
 
-We use a feed-forward neural network with an one fully connected layer, which predict the digit that has been in input image.
-For an image $x$, initialy $28 \times 28$, we flatten this array into a vector of $784 = 28 \times 28$.
+We use a feed-forward neural network with an one fully connected layer:
+\begin{align*}
+y &= f(Wx + b) \\
+label &= argmax(y)
+\end{align*}
 
-\begin{align}
-y &= softmax(Wx + b) \\
-label &= argmax(y)
-\end{align}
-
-In $(1)$, $W$ is the weights matrix, $10 \times 784$ -- the number of classes(one for each digit) and the dimension of an image ($784$).
+Where $W$ is the weights matrix, $10 \times 784$ -- the number of classes(one for each digit) and the dimension of an image ($784$).
 $b$ is the bias vector, $1 \times 10$, corresponding to the free element.
+$y$ is a vector of probabilities, from which argmax extracts the predicted labe
 \\
@@
 The activation function output a vector of probabilities:
-$$softmax(x_i)=\frac{exp(x_i)}{\sum_j exp(x_j)}$$
+$$f(x_i)=\frac{exp(x_i)}{\sum_j exp(x_j)}$$
 Where $exp(x)$ represents $e^x$, Euler's constant. It \textit{normalizes} the vector, causing the greatest elements to have logarithmic relative value.
 \\
-
-$(2)$ With $y$ vector of probabilities, $argmax$ take the index of the maximum value and results the prediction label.
 
@@
 
-\TODO{more about evaluation?}
-
-\noindent The following chart illustrates model evolution, both on the test and on the training set, after each pass through the dataset.
+\noindent The following chart illustrates model evolution, after each pass through the dataset, on the entire MNIST training (60 000) and test (10 000) dataset.
 
@@
   ylabel={Accuracy (\%)},
-  xmin=0, xmax=100,
+  xmin=0, xmax=250,
   ymin=0, ymax=100,
-  xtick={0,25,50,75,100},
+  enlarge x limits=0.05,
+  xtick={0,50,100,150,200, 250},
   ytick={0,25,50,75,100},
@@
 
+\addplot[dash pattern=on 6pt off 4pt]
+  coordinates {
+ (0, 25.542341)    (1, 47.745369)    (2, 66.106474)    (3, 79.268610)    (4, 81.409645)    (5, 82.697285)    (6, 83.599091)    (7, 84.318434)    (8, 84.974273)    (9, 85.386156)    (10, 85.663764)    (11, 86.105388)    (12, 86.463159)    (13, 86.705436)    (14, 86.973765)    (15, 87.296094)    (16, 87.391127)    (17, 87.672536)    (18, 87.834651)    (19, 87.989275)    (20, 88.216571)    (21, 88.300424)    (22, 88.423408)    (23, 88.635834)    (24, 88.689947)    (25, 88.783079)    (26, 88.917243)    (27, 89.086849)    (28, 89.001096)    (29, 89.200441)    (30, 89.248964)    (31, 89.334605)    (32, 89.433439)    (33, 89.565703)    (34, 89.670015)    (35, 89.733408)    (36, 89.835932)    (37, 89.940244)    (38, 89.940244)    (39, 90.016606)    (40, 90.037178)    (41, 90.124720)    (42, 90.079998)    (43, 90.175031)    (44, 90.219753)    (45, 90.365097)    (46, 90.448950)    (47, 90.372588)    (48, 90.560753)    (49, 90.534703)    (50, 90.560753)    (51, 90.687426)    (52, 90.603574)    (53, 90.763900)    (54, 90.821590)    (55, 90.810410)    (56, 90.843951)    (57, 90.920425)    (58, 90.901753)    (59, 90.914834)    (60, 90.948375)    (61, 91.037818)    (62, 91.007967)    (63, 91.136541)    (64, 91.125360)    (65, 91.220393)    (66, 91.147721)    (67, 91.233474)    (68, 91.328507)    (69, 91.334097)    (70, 91.384409)    (71, 91.255835)    (72, 91.330296)    (73, 91.438410)    (74, 91.348967)    (75, 91.430919)    (76, 91.522262)    (77, 91.542722)    (78, 91.546524)    (79, 91.531542)    (80, 91.596835)    (81, 91.555803)    (82, 91.624786)    (83, 91.643345)    (84, 91.783100)    (85, 91.768230)    (86, 91.712328)    (87, 91.727198)    (88, 91.818541)    (89, 91.840902)    (90, 91.755149)    (91, 91.900493)    (92, 91.846492)    (93, 91.870754)    (94, 91.909885)    (95, 91.971377)    (96, 91.909885)    (97, 92.034657)    (98, 91.904295)    (99, 91.973165)    (100, 91.982557)    (101, 92.010508)    (102, 92.006706)    (103, 92.012297)    (104, 92.083180)    (105, 92.045838)    (106, 92.131591)    (107, 92.213543)    (108, 92.072000)    (109, 92.189393)    (110, 92.146461)    (111, 92.254575)    (112, 92.245295)    (113, 92.185592)    (114, 92.293706)    (115, 92.273246)    (116, 92.258264)    (117, 92.381248)    (118, 92.278836)    (119, 92.474492)    (120, 92.358887)    (121, 92.375658)    (122, 92.476281)    (123, 92.448330)    (124, 92.351508)    (125, 92.511722)    (126, 92.425969)    (127, 92.388739)    (128, 92.360788)    (129, 92.450231)    (130, 92.502443)    (131, 92.565723)    (132, 92.543363)    (133, 92.621625)    (134, 92.647787)    (135, 92.532182)    (136, 92.560133)    (137, 92.614246)    (138, 92.629116)    (139, 92.578804)    (140, 92.696198)    (141, 92.703689)    (142, 92.586295)    (143, 92.714869)    (144, 92.727838)    (145, 92.657067)    (146, 92.737230)    (147, 92.714869)    (148, 92.660756)    (149, 92.711068)    (150, 92.726050)    (151, 92.735329)    (152, 92.843443)    (153, 92.789330)    (154, 92.828461)    (155, 92.742820)    (156, 92.766970)    (157, 92.876984)    (158, 92.875083)    (159, 92.722248)    (160, 92.794920)    (161, 92.871394)    (162, 92.826673)    (163, 92.878773)    (164, 92.938476)    (165, 92.882574)    (166, 92.958936)    (167, 92.882574)    (168, 92.942165)    (169, 92.895543)    (170, 92.785641)    (171, 92.988788)    (172, 92.953346)    (173, 92.899345)    (174, 92.951445)    (175, 93.014838)    (176, 93.026018)    (177, 92.981297)    (178, 93.018527)    (179, 92.975707)    (180, 93.016738)    (181, 93.024117)    (182, 93.024117)    (183, 93.009248)    (184, 92.984986)    (185, 93.085609)    (186, 92.975707)    (187, 93.052068)    (188, 93.095001)    (189, 93.027919)    (190, 93.083820)    (191, 93.087510)    (192, 92.975707)    (193, 93.063249)    (194, 93.020428)    (195, 92.930985)    (196, 93.135921)    (197, 93.100591)    (198, 93.111771)    (199, 93.115461)    (200, 93.130331)    (201, 93.206804)    (202, 93.121051)    (203, 93.122952)    (204, 93.063249)    (205, 93.171362)    (206, 93.271986)    (207, 93.076330)    (208, 93.156493)    (209, 93.204903)    (210, 93.225363)    (211, 93.197413)    (212, 93.162083)    (213, 93.221674)    (214, 93.242134)    (215, 93.141511)    (216, 93.227264)    (217, 93.201214)    (218, 93.247724)    (219, 93.271986)    (220, 93.234755)    (221, 93.245935)    (222, 93.208593)    (223, 93.307427)    (224, 93.262706)    (225, 93.271986)    (226, 93.298036)    (227, 93.275675)    (228, 93.260805)    (229, 93.130331)    (230, 93.184443)    (231, 93.251525)    (232, 93.214183)    (233, 93.296247)    (234, 93.283166)    (235, 93.296247)    (236, 93.251525)    (237, 93.318608)    (238, 93.324198)    (239, 93.314806)    (240, 93.385690)    (241, 93.305527)    (242, 93.333477)    (243, 93.335378)    (244, 93.255215)    (245, 93.337167)    (246, 93.329788)    (247, 93.286855)    (248, 93.372609)    (249, 93.331577)    (250, 93.314806)
+  };
 \addplot[]
   coordinates {
-  (0,25)(25,65)(50,90)(100,95)
-  };
-\addplot[dash pattern=on 6pt off 4pt]
-  coordinates {
-  (0,25)(25,50)(50,80)(100,90)
+(0, 24.432432)  (1, 44.549688)(2, 56.596855)(3, 62.102859)(4, 66.083265)(5, 68.563395)(6, 71.450773)(7, 70.905093)(8, 73.653296)(9, 74.766428)(10, 75.325960)(11, 76.262903)(12, 75.887467)(13, 77.394473)(14, 78.915509)(15, 78.533935)(16, 77.963218)(17, 78.724613)(18, 79.106624)(19, 79.873274)(20, 79.106624)(21, 79.297958)(22, 80.065484)(23, 80.643426)(24, 80.643426)(25, 79.681283)(26, 79.681283)(27, 79.489511)(28, 80.450560)(29, 81.029814)(30, 80.836510)(31, 81.611037)(32, 81.805215)(33, 81.417077)(34, 81.611037)(35, 81.611037)(36, 81.417077)(37, 82.779383)(38, 82.194226)(39, 82.194226)(40, 81.999611)(41, 82.584112)(42, 83.366506)(43, 82.779383)(44, 82.584112)(45, 83.955593)(46, 83.759013)(47, 83.759013)(48, 83.955593)(49, 84.744100)(50, 84.744100)(51, 83.759013)(52, 84.546646)(53, 83.759013)(54, 84.546646)(55, 84.941772)(56, 83.955593)(57, 83.759013)(58, 82.974872)(59, 84.152393)(60, 84.349410)(61, 84.744100)(62, 84.744100)(63, 85.139662)(64, 85.139662)(65, 85.734641)(66, 84.546646)(67, 85.933403)(68, 85.536096)(69, 85.139662)(70, 86.132384)(71, 85.734641)(72, 85.734641)(73, 86.730633)(74, 85.734641)(75, 85.139662)(76, 85.536096)(77, 85.734641)(78, 85.933403)(79, 86.331582)(80, 85.734641)(81, 85.139662)(82, 86.730633)(83, 85.139662)(84, 85.139662)(85, 86.930485)(86, 86.530999)(87, 86.530999)(88, 85.139662)(89, 86.132384)(90, 86.530999)(91, 86.530999)(92, 86.132384)(93, 86.530999)(94, 86.331582)(95, 86.530999)(96, 86.730633)(97, 86.132384)(98, 86.530999)(99, 87.130555)(100, 86.930485)(101, 86.331582)(102, 86.730633)(103, 86.530999)(104, 86.530999)(105, 85.734641)(106, 86.730633)(107, 86.530999)(108, 86.730633)(109, 85.337770)(110, 86.331582)(111, 85.734641)(112, 86.132384)(113, 86.530999)(114, 86.331582)(115, 86.331582)(116, 86.132384)(117, 86.331582)(118, 86.530999)(119, 86.930485)(120, 86.930485)(121, 85.933403)(122, 86.331582)(123, 86.930485)(124, 86.930485)(125, 86.530999)(126, 87.330843)(127, 87.130555)(128, 85.933403)(129, 86.132384)(130, 86.331582)(131, 85.933403)(132, 87.130555)(133, 86.530999)(134, 86.730633)(135, 86.331582)(136, 85.734641)(137, 86.730633)(138, 86.930485)(139, 86.930485)(140, 86.530999)(141, 87.531349)(142, 87.330843)(143, 86.331582)(144, 87.130555)(145, 86.530999)(146, 86.930485)(147, 87.531349)(148, 87.130555)(149, 86.530999)(150, 86.930485)(151, 87.933014)(152, 87.330843)(153, 87.330843)(154, 86.930485)(155, 86.930485)(156, 87.330843)(157, 87.130555)(158, 87.732072)(159, 87.330843)(160, 86.730633)(161, 87.531349)(162, 87.732072)(163, 86.930485)(164, 87.130555)(165, 87.933014)(166, 87.531349)(167, 87.330843)(168, 87.130555)(169, 87.933014)(170, 87.933014)(171, 87.330843)(172, 87.732072)(173, 87.531349)(174, 87.531349)(175, 87.531349)(176, 87.732072)(177, 87.732072)(178, 87.732072)(179, 87.732072)(180, 88.134173)(181, 87.531349)(182, 87.732072)(183, 87.933014)(184, 87.531349)(185, 87.531349)(186, 86.530999)(187, 88.537144)(188, 87.330843)(189, 86.930485)(190, 88.335550)(191, 87.330843)(192, 87.933014)(193, 87.130555)(194, 87.130555)(195, 87.330843)(196, 87.732072)(197, 88.134173)(198, 87.330843)(199, 88.134173)(200, 87.933014)(201, 87.130555)(202, 87.531349)(203, 86.930485)(204, 87.130555)(205, 87.330843)(206, 87.732072)(207, 87.933014)(208, 87.130555)(209, 87.933014)(210, 87.330843)(211, 87.933014)(212, 87.933014)(213, 87.732072)(214, 88.335550)(215, 87.531349)(216, 87.732072)(217, 87.732072)(218, 87.933014)(219, 88.134173)(220, 86.930485)(221, 87.933014)(222, 87.732072)(223, 87.732072)(224, 88.134173)(225, 87.933014)(226, 88.335550)(227, 88.134173)(228, 87.933014)(229, 88.537144)(230, 87.933014)(231, 87.531349)(232, 87.933014)(233, 87.933014)(234, 87.330843)(235, 87.330843)(236, 87.531349)(237, 87.933014)(238, 86.730633)(239, 87.933014)(240, 87.330843)(241, 87.933014)(242, 88.134173)(243, 87.933014)(244, 88.537144)(245, 87.933014)(246, 87.732072)(247, 87.933014)(248, 87.531349)(249, 88.537144)(250, 88.335550)
   };
@@
 
-\TODO{comment on chart}
+The training performance is intentionally dashed, as  what interests us is the model's generalization power.
+It quickly learns in the first 50 epochs and reaches a plateau around the 100 mark.
+It achieves 93.3\% performance on the training set and 88.3\% on unseen data.
 

### 2017-02-09 doc/Paper.tex
add interface ss to paper

@@
 \usepackage{todonotes}
+\usepackage{wrapfig}
 
@@
 \section{Graphical Interface}
+
 The user is presented with a blank box in which to draw a digit to be recognized. Upon request, the prediction along with the model's confidence in it is displayed.
 
-\missingfigure[figcolor=white]{GUI}
+\begin{wrapfigure}{r}{0.6\textwidth}\centering
+    \includegraphics[width=6cm]{images/gui.png}
+\end{wrapfigure}
 
@@
 
-\missingfigure[figcolor=white]{resizing example pdf}
+\begin{center}
+\includegraphics[width=\textwidth]{images/resize.png}
+\end{center}
+
 The transformation can be computed by \textit{compacting} the columns:

### 2017-02-09 main.cpp

@@ -40,7 +40,7 @@ void predict(string pixels_str) {
     parse_image(pixels_str, input);
 
     // Read
-    NeuralNetwork net("models/trained.nn");     // TODO: names as additional args
+    NeuralNetwork net("models/best.nn");     // TODO: names as additional args
 
     // Predict
     auto prediction = net.predict_one(input);
@@ -51,8 +51,8 @@ int main(int argc, const char* argv[]) {
 //    string dummy = "0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.03973336818716507,0.36884067442164425,0.3999477192523853,0.40007842112142206,0.292772186642269,0.06665795320873089,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.04339302052019344,0.7137629068095673,0.9998692981309633,0.9998692981309633,0.9998692981309633,0.9998692981309633,0.9601359299437981,0.3708012024571951,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.47497059207946674,0.9998692981309633,0.9998692981309633,0.9998692981309633,0.9997385962619265,0.9998692981309633,0.9998692981309633,0.9823552476800418,0.07554568030322834,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.14560188210691413,0.9231473010064044,0.9998692981309633,0.9334727486603058,0.012808783165599268,0.013462292510782905,0.37746699777806825,0.9998692981309633,0.9998692981309633,0.13331590641746177,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.009018428963534178,0.7947980656123382,0.9998692981309633,0.9861456018821069,0.46974251731799765,0,0,0.26663181283492354,0.9998692981309633,0.9998692981309633,0.13331590641746177,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3423081950071886,0.9998692981309633,0.9998692981309633,0.425565285
[…]

### 2017-02-09 doc/Paper.tex
amend

@@
 \begin{abstract}
-\TODO{rewrite when paper is done, talk about each section}
-
-We implement a neural network to classify handwritten digits.
-The training is done in parallel, using MPI.
-We provide a comparison for model hyper-parameters and for training time vs number of processes.
-It achieves great accuracy and fast training time.
-An interactive demo is also presented
+We implement a neural network to classify handwritten digits with training done in parallel, using MPI.
+We present the dataset, model architecture, parallelization strategy and an interactive interface.
+Then we discuss model accuracy evolution and time performance benchmarks.
+Our implementation achieves good prediction performance with fast training time.
 \end{abstract}
@@
 
-We will present the details, data used, model architecture, parallelization strategy and an interactive interface.
+In the following sections we will talk about the dataset used; detail model architecture and reason for choosing it along with its training performance.
+Parallelism has a dedicated section with a high-level overview of the solution and time performance benchmarks.
+The last sections contain an outline of the graphical interface and implementation details for the entire project.
 
@@
 $$\frac{1}{\lambda} \sum_i \sum_j W_{i,j}^2$$
-Where $\lambda$ is the regularization parameter. This limits overfitting by imposing a penalty on big weights.
+Where $\lambda$ is the regularization parameter. This limits over-fitting by imposing a penalty on big weights.
 \\
